\documentclass[11pt]{article}
\input{_preamble}
\fancyhead[R]{\footnotesize MBF3C \textperiodcentered{} 5.2}
\begin{document}

\begin{center}
{\LARGE\bfseries\color{titleblue}Worksheet 5.2 --- Central Tendency \& Spread}\\[2pt]
{\small Grade 11 Foundations for College Mathematics (MBF3C) \textperiodcentered{} Unit 5: Data Management}
\end{center}

\begin{ideabox}
Summarize data with a centre (mean, median, mode) and a spread (range, standard deviation); choose the measure that fits the data.
\begin{itemize}[leftmargin=*,itemsep=2pt]
  \item Mean $=$ sum$\div$count; median $=$ middle (average two middles if even); mode $=$ most frequent.
  \item Range $=$ max $-$ min.
  \item The median resists outliers that distort the mean.
\end{itemize}
\end{ideabox}

\begin{learnbox}
\lh{Measures of centre}The \textbf{mean} is the average, the \textbf{median} is the middle value, and the \textbf{mode} is the most common value.
\lh{Mean vs median}The mean uses every value but is pulled by outliers; the median resists them, so it suits skewed data.
\lh{Spread}The \textbf{range} — highest minus lowest — shows how spread out the data is.
\end{learnbox}

\sech{Worked Examples}

\begin{example}{ex0}{Example 1 --- Mean}
Mean of $4,8,6,10,2$.\soln Sum $=30$, count $=5$, mean $=6$.
\end{example}

\begin{example}{ex1}{Example 2 --- Median (even count)}
Median of $3,7,8,12$.\soln Even count: average the middles, $\dfrac{7+8}{2}=7.5$.
\end{example}

\begin{example}{ex2}{Example 3 --- Mode and range}
For $5,5,7,9,9,9,12$: mode and range.\soln Mode $=9$; range $=12-5=7$.
\end{example}

\begin{example}{ex3}{Example 4 --- Outlier: mean vs median}
For $2,3,4,5,100$, find both; which is typical?\soln Mean $=\dfrac{114}{5}=22.8$; median $=4$. The median is more typical — it resists the outlier $100$.
\end{example}

\begin{example}{ex4}{Example 5 --- Choose the measure}
(a) favourite flavour, (b) incomes (a few very high), (c) symmetric scores.\soln (a) mode, (b) median, (c) mean.
\end{example}

\begin{example}{ex5}{Example 6 --- Mean}
Mean of $10,12,14,16,18$.\soln $70\div5=14$.
\end{example}

\begin{example}{ex6}{Example 7 --- Median (even count)}
Median of $2,5,9,11,14,20$.\soln $\dfrac{9+11}{2}=10$.
\end{example}

\begin{example}{ex7}{Example 8 --- Mode and range}
For $4,4,4,7,9,9,15$: mode and range.\soln Mode $=4$; range $=11$.
\end{example}

\begin{example}{ex8}{Example 9 --- Compare spread}
Two classes share a mean of $70$ but ranges $10$ and $40$. What does that show?\soln Same centre, but the second class is far more spread out.
\end{example}

\clearpage
\sech{Your Turn --- Show Your Work}
For each question, show all steps clearly in the space provided.

\begin{qbox}{Question 1}
Mean of $10,12,14,16,18$?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 2}
Median of $2,5,9,11,14,20$?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 3}
Mode and range of $4,4,4,7,9,9,15$?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 4}
For $1,2,2,3,40$, is the mean or median more typical?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 5}
Mean of $3,7,7,9,14$?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 6}
Median of $6,2,9,4,8$ (order first)?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 7}
Range of $15,8,23,4,19$?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 8}
Mode of $2,4,4,4,6,6,9$?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 9}
A data set has mean $50$; adding the value $200$ will do what to the mean?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 10}
Median of $4,4,6,10$?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 11}
Which measure suits "most common shoe size sold"?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 12}
Mean of $0,0,0,20$?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 13 --- Challenge}
Test scores: $60,65,70,70,75,90$. Find the mean, median, and mode, and say which best summarizes them.\workspace{3cm}
\end{qbox}

\clearpage
\sech{Question Recap \& Answer Key}

\begin{recapbox}
\begin{multicols}{2}
\small
\begin{enumerate}[leftmargin=*,itemsep=3pt]
  \item Mean of $10,12,14,16,18$?
  \item Median of $2,5,9,11,14,20$?
  \item Mode and range of $4,4,4,7,9,9,15$?
  \item For $1,2,2,3,40$, is the mean or median more typical?
  \item Mean of $3,7,7,9,14$?
  \item Median of $6,2,9,4,8$ (order first)?
  \item Range of $15,8,23,4,19$?
  \item Mode of $2,4,4,4,6,6,9$?
  \item A data set has mean $50$; adding the value $200$ will do what to the mean?
  \item Median of $4,4,6,10$?
  \item Which measure suits "most common shoe size sold"?
  \item Mean of $0,0,0,20$?
  \item Test scores: $60,65,70,70,75,90$. Find the mean, median, and mode, and say which best summarizes them.
\end{enumerate}
\end{multicols}
\end{recapbox}

\begin{keybox}
\begin{multicols}{2}
\small
\begin{enumerate}[leftmargin=*,itemsep=3pt]
  \item $14$
  \item $10$
  \item mode $4$, range $11$
  \item median ($=2$; mean $=9.6$ is distorted)
  \item $8$
  \item $2,4,6,8,9\Rightarrow6$
  \item $23-4=19$
  \item $4$
  \item raise it (outlier pulls the mean up)
  \item $\dfrac{4+6}{2}=5$
  \item mode
  \item $5$
  \item mean $=71.7$, median $=70$, mode $=70$; median/mode best (slight high outlier 90)
\end{enumerate}
\end{multicols}
\end{keybox}

\end{document}
