\documentclass[11pt]{article}
\input{_preamble}
\fancyhead[R]{\footnotesize MHF4U \textperiodcentered{} 1.2}
\begin{document}

\begin{center}
{\LARGE\bfseries\color{titleblue}Worksheet 1.2 --- Characteristics of Polynomial Functions}\\[2pt]
{\small Grade 12 Advanced Functions (MHF4U) \textperiodcentered{} Unit 1: Polynomial Functions}
\end{center}

\begin{ideabox}
Degree $n$ limits the graph: at most $n$ zeros and $n-1$ turning points; the $n$-th finite differences are constant.
\begin{itemize}[leftmargin=*,itemsep=2pt]
  \item At most $n$ x-intercepts and $n-1$ turning points.
  \item The $n$-th differences of a degree-$n$ polynomial are constant.
  \item Even: $f(-x)=f(x)$ (y-axis). Odd: $f(-x)=-f(x)$ (origin).
\end{itemize}
\end{ideabox}

\begin{learnbox}
\lh{Built from power functions}A \textbf{polynomial} is a sum of power functions; its degree still governs end behaviour and caps the number of intercepts and turning points.
\lh{Key features}Read a polynomial by its degree, leading coefficient, zeros (with multiplicity), and turning points — at most $n-1$ for degree $n$.
\lh{Finite differences}The degree shows up in a table too: the $n$th finite differences of a degree-$n$ polynomial are constant.
\end{learnbox}

\sech{Worked Examples}

\begin{example}{ex0}{Example 1 --- Turning points}
Max turning points of a degree-5 polynomial?\soln $n-1=5-1=4$.
\end{example}

\begin{example}{ex1}{Example 2 --- x-intercepts}
Max x-intercepts of a degree-4 polynomial?\soln At most $n=4$.
\end{example}

\begin{example}{ex2}{Example 3 --- Degree from differences}
The 3rd finite differences are constant. Degree?\soln $n$-th differences constant ⇒ degree $n$, so $3$.
\end{example}

\begin{example}{ex3}{Example 4 --- Even function}
Classify $f(x)=x^4-3x^2+1$.\soln Only even powers ⇒ $f(-x)=f(x)$: even (y-axis symmetry).
\end{example}

\begin{example}{ex4}{Example 5 --- Odd function}
Classify $f(x)=x^3-4x$.\soln Only odd powers ⇒ $f(-x)=-f(x)$: odd (origin symmetry).
\end{example}

\begin{example}{ex5}{Example 6 --- A cubic's graph}
A cubic $y=x^3-3x$ has how many turning points and zeros?\soln 2 turning points (at $x=\pm1$) and 3 zeros — the maxima for degree 3:\eplot{-3}{3}{-4}{4}{\addplot[exblue,very thick,domain=-2.1:2.1]{x^3-3*x};\addplot[red,only marks,mark=*,mark size=1.6pt] coordinates {(-1,2) (1,-2)};}
\end{example}

\begin{example}{ex6}{Example 7 --- Minimum degree}
A polynomial has 4 turning points. Minimum degree?\soln Turning points $\le n-1$, so $n\ge5$.
\end{example}

\begin{example}{ex7}{Example 8 --- Degree from differences}
The 2nd differences are constant. Degree?\soln $2$.
\end{example}

\begin{example}{ex8}{Example 9 --- Neither}
Classify $f(x)=x^3+x^2$.\soln Mixed even and odd powers ⇒ neither.
\end{example}

\clearpage
\sech{Your Turn --- Show Your Work}
For each question, show all steps clearly in the space provided.

\begin{qbox}{Question 1}
Max turning points of a degree-4 polynomial?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 2}
Max x-intercepts of a degree-6 polynomial?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 3}
The 4th differences are constant. Degree?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 4}
Is $f(x)=x^6-x^2$ even, odd, or neither?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 5}
Is $f(x)=x^5-x^3$ even, odd, or neither?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 6}
Max turning points of a degree-7 polynomial?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 7}
A polynomial has 5 turning points. Minimum degree?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 8}
The 1st differences are constant. Degree?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 9}
Is $f(x)=x^4+x$ even, odd, or neither?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 10}
Max x-intercepts of a degree-3 polynomial?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 11}
Symmetry of an even function?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 12}
Symmetry of an odd function?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 13 --- Challenge}
A degree-4 polynomial has exactly 1 turning point — is that possible? Explain.\workspace{3cm}
\end{qbox}

\clearpage
\sech{Question Recap \& Answer Key}

\begin{recapbox}
\begin{multicols}{2}
\small
\begin{enumerate}[leftmargin=*,itemsep=3pt]
  \item Max turning points of a degree-4 polynomial?
  \item Max x-intercepts of a degree-6 polynomial?
  \item The 4th differences are constant. Degree?
  \item Is $f(x)=x^6-x^2$ even, odd, or neither?
  \item Is $f(x)=x^5-x^3$ even, odd, or neither?
  \item Max turning points of a degree-7 polynomial?
  \item A polynomial has 5 turning points. Minimum degree?
  \item The 1st differences are constant. Degree?
  \item Is $f(x)=x^4+x$ even, odd, or neither?
  \item Max x-intercepts of a degree-3 polynomial?
  \item Symmetry of an even function?
  \item Symmetry of an odd function?
  \item A degree-4 polynomial has exactly 1 turning point — is that possible? Explain.
\end{enumerate}
\end{multicols}
\end{recapbox}

\begin{keybox}
\begin{multicols}{2}
\small
\begin{enumerate}[leftmargin=*,itemsep=3pt]
  \item 3
  \item 6
  \item 4
  \item even
  \item odd
  \item 6
  \item 6
  \item 1
  \item neither
  \item 3
  \item y-axis
  \item origin
  \item yes (it need not reach the maximum of 3)
\end{enumerate}
\end{multicols}
\end{keybox}

\end{document}
