\documentclass[11pt]{article}
\input{_preamble}
\fancyhead[R]{\footnotesize MHF4U \textperiodcentered{} 1.4}
\begin{document}

\begin{center}
{\LARGE\bfseries\color{titleblue}Worksheet 1.4 --- Transformations of Functions}\\[2pt]
{\small Grade 12 Advanced Functions (MHF4U) \textperiodcentered{} Unit 1: Polynomial Functions}
\end{center}

\begin{ideabox}
Apply $g(x)=a\,f\big(k(x-d)\big)+c$ to the parent $f(x)=x^n$.
\begin{itemize}[leftmargin=*,itemsep=2pt]
  \item $a$: vertical stretch/reflection. $k$: horizontal stretch/reflection.
  \item $d$: horizontal shift (opposite to the sign). $c$: vertical shift.
  \item Stretches/reflections first, then translations.
\end{itemize}
\end{ideabox}

\begin{learnbox}
\lh{A single framework}Any base function $f$ transforms through $y=af\!\left(k(x-d)\right)+c$, where each constant stretches, reflects, or shifts the graph.
\lh{Vertical vs horizontal}$a$ and $c$ act vertically (stretch by $a$, shift by $c$); $k$ and $d$ act horizontally (compress by $k$, shift by $d$).
\lh{Order and mapping}Apply horizontal changes to $x$ and vertical changes to $y$: each point maps as $(x,y)\to\left(\tfrac{x}{k}+d,\ ay+c\right)$.
\end{learnbox}

\sech{Worked Examples}

\begin{example}{ex0}{Example 1 --- Describe}
Describe $y=2(x-1)^3+3$ from $y=x^3$.\soln $a=2$ stretch; $d=1$ right; $c=3$ up — centred at $(1,3)$:\eplot{-2}{4}{-6}{12}{\addplot[exblue,very thick,domain=-0.4:2.4]{2*(x-1)^3+3};\addplot[red,only marks,mark=*,mark size=1.6pt] coordinates {(1,3)};}
\end{example}

\begin{example}{ex1}{Example 2 --- Reflection}
Describe $y=-x^4$.\soln $a=-1$: reflection in the x-axis (opens down).
\end{example}

\begin{example}{ex2}{Example 3 --- Horizontal stretch}
What does $k=2$ do in $y=(2x)^3$?\soln Horizontal compression by $\tfrac12$ (squeezed toward the y-axis).
\end{example}

\begin{example}{ex3}{Example 4 --- Map a point}
$(1,1)$ on $y=x^3$ maps where on $y=(x-2)^3+4$?\soln Right 2, up 4: $(3,5)$.
\end{example}

\begin{example}{ex4}{Example 5 --- Write the equation}
Write $y=x^4$ reflected, left 1, down 2.\soln $a=-1,\ d=-1,\ c=-2$: $y=-(x+1)^4-2$.
\end{example}

\begin{example}{ex5}{Example 6 --- Describe}
Describe $y=(x+3)^3-1$.\soln Left 3, down 1.
\end{example}

\begin{example}{ex6}{Example 7 --- Map a point}
$(1,1)$ on $y=x^3$ maps where on $y=x^3-5$?\soln $(1,-4)$.
\end{example}

\begin{example}{ex7}{Example 8 --- Stretch \& reflect}
Describe $y=-3x^4+2$.\soln Reflect, stretch by 3, up 2.
\end{example}

\begin{example}{ex8}{Example 9 --- Write the equation}
Write $y=x^3$ shifted right 4 and up 1.\soln $y=(x-4)^3+1$.
\end{example}

\clearpage
\sech{Your Turn --- Show Your Work}
For each question, show all steps clearly in the space provided.

\begin{qbox}{Question 1}
Describe $y=(x+3)^3-1$ from $y=x^3$.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 2}
What does $a=-2$ do?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 3}
Where does $(1,1)$ on $y=x^3$ map under $y=x^3-5$?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 4}
Write $y=x^3$ shifted right 4 and up 1.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 5}
Describe $y=-3x^4+2$.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 6}
Write $y=x^4$ reflected, left 1 and down 2.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 7}
Where does $(1,1)$ on $y=x^3$ map under $y=(x-2)^3+4$?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 8}
What does $k=3$ do?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 9}
Describe $y=2(x+1)^4$ from $y=x^4$.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 10}
Write $y=x^3$ reflected in the x-axis.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 11}
State the vertex-like centre of $y=(x-5)^3+2$.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 12}
Describe $y=\tfrac12 x^3$.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 13 --- Challenge}
Fully describe $y=-2(x+1)^3-3$.\workspace{3cm}
\end{qbox}

\clearpage
\sech{Question Recap \& Answer Key}

\begin{recapbox}
\begin{multicols}{2}
\small
\begin{enumerate}[leftmargin=*,itemsep=3pt]
  \item Describe $y=(x+3)^3-1$ from $y=x^3$.
  \item What does $a=-2$ do?
  \item Where does $(1,1)$ on $y=x^3$ map under $y=x^3-5$?
  \item Write $y=x^3$ shifted right 4 and up 1.
  \item Describe $y=-3x^4+2$.
  \item Write $y=x^4$ reflected, left 1 and down 2.
  \item Where does $(1,1)$ on $y=x^3$ map under $y=(x-2)^3+4$?
  \item What does $k=3$ do?
  \item Describe $y=2(x+1)^4$ from $y=x^4$.
  \item Write $y=x^3$ reflected in the x-axis.
  \item State the vertex-like centre of $y=(x-5)^3+2$.
  \item Describe $y=\tfrac12 x^3$.
  \item Fully describe $y=-2(x+1)^3-3$.
\end{enumerate}
\end{multicols}
\end{recapbox}

\begin{keybox}
\begin{multicols}{2}
\small
\begin{enumerate}[leftmargin=*,itemsep=3pt]
  \item left 3, down 1
  \item reflect and stretch by 2
  \item $(1,-4)$
  \item $y=(x-4)^3+1$
  \item reflect, stretch 3, up 2
  \item $y=-(x+1)^4-2$
  \item $(3,5)$
  \item horizontal compression by $\tfrac13$
  \item stretch 2, left 1
  \item $y=-x^3$
  \item $(5,2)$
  \item vertical compression by $\tfrac12$
  \item reflect, stretch 2, left 1, down 3
\end{enumerate}
\end{multicols}
\end{keybox}

\end{document}
