\documentclass[11pt]{article}
\input{_preamble}
\fancyhead[R]{\footnotesize MHF4U \textperiodcentered{} 2.1}
\begin{document}

\begin{center}
{\LARGE\bfseries\color{titleblue}Worksheet 2.1 --- Dividing Polynomials}\\[2pt]
{\small Grade 12 Advanced Functions (MHF4U) \textperiodcentered{} Unit 2: Polynomial Equations and Inequalities}
\end{center}

\begin{ideabox}
Long division always works; synthetic division is a shortcut for dividing by $x-a$. Every result is a division statement $P(x)=D(x)Q(x)+R(x)$.
\begin{itemize}[leftmargin=*,itemsep=2pt]
  \item Long division: divide, multiply, subtract, bring down.
  \item Synthetic (by $x-a$): bring down, multiply by $a$, add.
  \item $\deg R<\deg D$; remainder $0$ means $D$ is a factor.
\end{itemize}
\end{ideabox}

\begin{learnbox}
\lh{The division identity}Polynomial division gives $P(x)=D(x)Q(x)+R$, with the remainder $R$ lower in degree than the divisor $D$.
\lh{Long and synthetic}Long division works for any divisor; \textbf{synthetic division} is a fast shortcut when the divisor is $x-a$.
\lh{Fill the gaps}Insert zero coefficients for missing powers so the terms line up correctly through the division.
\end{learnbox}

\sech{Worked Examples}

\begin{example}{ex0}{Example 1 --- Exact long division}
$(x^2+5x+6)\div(x+2)$.\soln $x^2\div x=x$; subtract $x^2+2x$ → $3x+6$; $3x\div x=3$; subtract → $0$. Quotient $x+3$, R $0$. The graph confirms a zero at $x=-2$:\eplot{-5}{1}{-1}{8}{\addplot[exblue,very thick,domain=-4.6:0.6]{x^2+5*x+6};\addplot[red,only marks,mark=*,mark size=1.6pt] coordinates {(-2,0) (-3,0)};}
\end{example}

\begin{example}{ex1}{Example 2 --- Exact long division}
$(x^2+7x+10)\div(x+5)$.\soln Quotient $x+2$, R $0$, so $x^2+7x+10=(x+5)(x+2)$.
\end{example}

\begin{example}{ex2}{Example 3 --- Synthetic division}
$(x^3-4x^2+x+6)\div(x-2)$.\soln Coefficients $1,-4,1,6$, root $2$: bring down $1$; $1{\cdot}2=2,\ -4{+}2=-2$; $-2{\cdot}2=-4,\ 1{-}4=-3$; $-3{\cdot}2=-6,\ 6{-}6=0$. Quotient $x^2-2x-3$, R $0$.
\end{example}

\begin{example}{ex3}{Example 4 --- Division statement}
Write the statement for $(x^2+5x+6)\div(x+2)$.\soln $x^2+5x+6=(x+2)(x+3)+0$.
\end{example}

\begin{example}{ex4}{Example 5 --- Nonzero remainder}
$(x^2+3x+5)\div(x+1)$.\soln Quotient $x+2$; $(x+1)(x+2)=x^2+3x+2$; subtract → $3$. So $x^2+3x+5=(x+1)(x+2)+3$.
\end{example}

\begin{example}{ex5}{Example 6 --- Synthetic}
$(x^3-1)\div(x-1)$.\soln Coefficients $1,0,0,-1$, root $1$: $1,1,1,0$. Quotient $x^2+x+1$, R $0$.
\end{example}

\begin{example}{ex6}{Example 7 --- Linear over linear}
$(x^2+x-6)\div(x-2)$.\soln Quotient $x+3$, R $0$, so $(x-2)(x+3)$.
\end{example}

\begin{example}{ex7}{Example 8 --- Remainder}
$(x^2+4x+1)\div(x+1)$.\soln Quotient $x+3$, remainder $-2$: $x^2+4x+1=(x+1)(x+3)-2$.
\end{example}

\begin{example}{ex8}{Example 9 --- Cubic remainder}
$(x^3+2x^2-5)\div(x-1)$.\soln Synthetic with root $1$ on $1,2,0,-5$: $1,3,3,-2$. Quotient $x^2+3x+3$, R $-2$.
\end{example}

\clearpage
\sech{Your Turn --- Show Your Work}
For each question, show all steps clearly in the space provided.

\begin{qbox}{Question 1}
$(x^2+6x+8)\div(x+2)$?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 2}
$(x^2+x-6)\div(x-2)$?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 3}
Synthetic: $(x^3-1)\div(x-1)$?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 4}
$(x^2+4x+1)\div(x+1)$, with remainder?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 5}
Write the division statement for $(x^2+4x+1)\div(x+1)$.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 6}
$(x^2+9x+20)\div(x+4)$?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 7}
$(x^3-8)\div(x-2)$?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 8}
$(x^2-5x+6)\div(x-3)$?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 9}
$(x^3+2x^2-5)\div(x-1)$, quotient and remainder?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 10}
Is $x+3$ a factor of $x^2+5x+6$? (use the remainder)\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 11}
$(2x^2+7x+3)\div(x+3)$?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 12}
$(x^2+1)\div(x+1)$, remainder?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 13 --- Challenge}
Divide $(x^3-2x^2-5x+6)\div(x-1)$ and state quotient and remainder.\workspace{3.5cm}
\end{qbox}

\clearpage
\sech{Question Recap \& Answer Key}

\begin{recapbox}
\begin{multicols}{2}
\small
\begin{enumerate}[leftmargin=*,itemsep=3pt]
  \item $(x^2+6x+8)\div(x+2)$?
  \item $(x^2+x-6)\div(x-2)$?
  \item Synthetic: $(x^3-1)\div(x-1)$?
  \item $(x^2+4x+1)\div(x+1)$, with remainder?
  \item Write the division statement for $(x^2+4x+1)\div(x+1)$.
  \item $(x^2+9x+20)\div(x+4)$?
  \item $(x^3-8)\div(x-2)$?
  \item $(x^2-5x+6)\div(x-3)$?
  \item $(x^3+2x^2-5)\div(x-1)$, quotient and remainder?
  \item Is $x+3$ a factor of $x^2+5x+6$? (use the remainder)
  \item $(2x^2+7x+3)\div(x+3)$?
  \item $(x^2+1)\div(x+1)$, remainder?
  \item Divide $(x^3-2x^2-5x+6)\div(x-1)$ and state quotient and remainder.
\end{enumerate}
\end{multicols}
\end{recapbox}

\begin{keybox}
\begin{multicols}{2}
\small
\begin{enumerate}[leftmargin=*,itemsep=3pt]
  \item $x+4$, R 0
  \item $x+3$, R 0
  \item $x^2+x+1$, R 0
  \item $x+3$, R $-2$
  \item $(x+1)(x+3)-2$
  \item $x+5$, R 0
  \item $x^2+2x+4$, R 0
  \item $x-2$, R 0
  \item $x^2+3x+3$, R $-2$
  \item yes (R 0)
  \item $2x+1$, R 0
  \item $x-1$, R 2
  \item $x^2-x-6$, R 0
\end{enumerate}
\end{multicols}
\end{keybox}

\end{document}
