\documentclass[11pt]{article}
\input{_preamble}
\fancyhead[R]{\footnotesize MHF4U \textperiodcentered{} 4.4}
\begin{document}

\begin{center}
{\LARGE\bfseries\color{titleblue}Worksheet 4.4 --- Applications of Exponential \& Log Models}\\[2pt]
{\small Grade 12 Advanced Functions (MHF4U) \textperiodcentered{} Unit 4: Exponential and Logarithmic Functions}
\end{center}

\begin{ideabox}
Model with $A=A_0\,b^{t/p}$ (growth $b>1$, decay $0<b<1$); use logs for doubling time and half-life; log scales (pH, dB, Richter) step by factors of ten.
\begin{itemize}[leftmargin=*,itemsep=2pt]
  \item $A=A_0\,b^{t/p}$, where $p$ is the doubling time or half-life.
  \item Doubling time $=\dfrac{\log2}{\log(\text{growth factor})}$.
  \item $\text{pH}=-\log[\text{H}^+]$; each Richter step is $\times10$.
\end{itemize}
\end{ideabox}

\begin{learnbox}
\lh{Growth and decay}Exponential models $A=A_0 b^{\,t}$ describe populations, radioactive decay, and cooling; logarithms solve them for the time.
\lh{Compound interest}Investments grow by $A=P(1+i)^n$, and finding the number of periods $n$ requires a logarithm.
\lh{Logarithmic scales}Quantities spanning huge ranges — sound in decibels, earthquakes on the Richter scale, acidity as pH — use \textbf{logarithmic} scales.
\end{learnbox}

\sech{Worked Examples}

\begin{example}{ex0}{Example 1 --- Growth}
$P=100\cdot2^{t/5}$. Find $P$ at $t=10$.\soln $P=100\cdot2^{2}=400$. Growth $2^x$ vs decay $0.5^x$:\eplot{-3}{3}{-1}{9}{\addplot[exblue,very thick,domain=-3:3.1,samples=80]{2^x};\addplot[qorange,very thick,domain=-3.1:3,samples=80]{0.5^x};}
\end{example}

\begin{example}{ex1}{Example 2 --- Half-life}
$A=80(\tfrac12)^{t/3}$. Find $A$ at $t=6$.\soln $A=80\cdot(\tfrac12)^2=80\cdot\tfrac14=20$ mg.
\end{example}

\begin{example}{ex2}{Example 3 --- Doubling time}
How long to double at 5\% /yr?\soln $2=1.05^t\Rightarrow t=\dfrac{\log2}{\log1.05}\approx14.2$ yr.
\end{example}

\begin{example}{ex3}{Example 4 --- pH scale}
Find the pH if $[\text{H}^+]=10^{-4}$.\soln $\text{pH}=-\log(10^{-4})=4$.
\end{example}

\begin{example}{ex4}{Example 5 --- Richter scale}
How much stronger is magnitude 7 than magnitude 5?\soln Difference $2\Rightarrow10^2=100\times$.
\end{example}

\begin{example}{ex5}{Example 6 --- Growth}
$P=200\cdot2^{t/4}$. Find $P$ at $t=8$.\soln $200\cdot2^2=800$.
\end{example}

\begin{example}{ex6}{Example 7 --- Half-life}
$A=160(\tfrac12)^{t/5}$. Find $A$ at $t=10$.\soln $160\cdot\tfrac14=40$.
\end{example}

\begin{example}{ex7}{Example 8 --- Doubling time}
How long to double at 10\% /yr?\soln $\dfrac{\log2}{\log1.1}\approx7.3$ yr.
\end{example}

\begin{example}{ex8}{Example 9 --- pH scale}
Find the pH if $[\text{H}^+]=10^{-7}$.\soln pH $=7$.
\end{example}

\clearpage
\sech{Your Turn --- Show Your Work}
For each question, show all steps clearly in the space provided.

\begin{qbox}{Question 1}
$P=300\cdot2^{t/6}$. Find $P$ at $t=12$.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 2}
$A=240(\tfrac12)^{t/4}$. Find $A$ at $t=8$.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 3}
How long to double at 8\% /yr (2 d.p.)?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 4}
pH if $[\text{H}^+]=10^{-3}$?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 5}
How much stronger is magnitude 6 than magnitude 3?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 6}
$P=100\cdot2^{t/10}$. Find $P$ at $t=20$.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 7}
$A=64(\tfrac12)^{t/2}$. Find $A$ at $t=6$.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 8}
How long to triple at 5\% /yr (2 d.p.)?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 9}
pH if $[\text{H}^+]=10^{-9}$?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 10}
A quake is $10^4$ times stronger than another. What is the magnitude difference?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 11}
$P=50\cdot3^{t/5}$. Find $P$ at $t=5$.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 12}
How long for 5\% /yr growth to reach 1.5 times the start (2 d.p.)?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 13 --- Challenge}
Bacteria double every 3 h from 500. Write the model and find the count at $t=9$ h.\workspace{3.5cm}
\end{qbox}

\clearpage
\sech{Question Recap \& Answer Key}

\begin{recapbox}
\begin{multicols}{2}
\small
\begin{enumerate}[leftmargin=*,itemsep=3pt]
  \item $P=300\cdot2^{t/6}$. Find $P$ at $t=12$.
  \item $A=240(\tfrac12)^{t/4}$. Find $A$ at $t=8$.
  \item How long to double at 8\% /yr (2 d.p.)?
  \item pH if $[\text{H}^+]=10^{-3}$?
  \item How much stronger is magnitude 6 than magnitude 3?
  \item $P=100\cdot2^{t/10}$. Find $P$ at $t=20$.
  \item $A=64(\tfrac12)^{t/2}$. Find $A$ at $t=6$.
  \item How long to triple at 5\% /yr (2 d.p.)?
  \item pH if $[\text{H}^+]=10^{-9}$?
  \item A quake is $10^4$ times stronger than another. What is the magnitude difference?
  \item $P=50\cdot3^{t/5}$. Find $P$ at $t=5$.
  \item How long for 5\% /yr growth to reach 1.5 times the start (2 d.p.)?
  \item Bacteria double every 3 h from 500. Write the model and find the count at $t=9$ h.
\end{enumerate}
\end{multicols}
\end{recapbox}

\begin{keybox}
\begin{multicols}{2}
\small
\begin{enumerate}[leftmargin=*,itemsep=3pt]
  \item $1200$
  \item $60$
  \item $\tfrac{\log2}{\log1.08}\approx9.01$ yr
  \item $3$
  \item $1000\times$
  \item $400$
  \item $8$
  \item $\tfrac{\log3}{\log1.05}\approx22.52$ yr
  \item $9$
  \item $4$
  \item $150$
  \item $\tfrac{\log1.5}{\log1.05}\approx8.31$ yr
  \item $P=500\cdot2^{t/3}$, $P(9)=4000$
\end{enumerate}
\end{multicols}
\end{keybox}

\end{document}
