\documentclass[11pt]{article}
\input{_preamble}
\fancyhead[R]{\footnotesize MHF4U \textperiodcentered{} 7.3}
\begin{document}

\begin{center}
{\LARGE\bfseries\color{titleblue}Worksheet 7.3 --- Composition of Functions}\\[2pt]
{\small Grade 12 Advanced Functions (MHF4U) \textperiodcentered{} Unit 7: Rates of Change and Combining Functions}
\end{center}

\begin{ideabox}
A composition feeds one function into another: $(f\circ g)(x)=f(g(x))$ — inner first, then outer. Order matters, and you can also decompose.
\begin{itemize}[leftmargin=*,itemsep=2pt]
  \item $(f\circ g)(x)=f(g(x))$: apply $g$ first, then $f$.
  \item Domain: $x$ in $g$'s domain and $g(x)$ in $f$'s domain.
  \item Decompose: split into an outer and inner function.
\end{itemize}
\end{ideabox}

\begin{learnbox}
\lh{One function inside another}\textbf{Composition} $(f\circ g)(x)=f(g(x))$ feeds the output of $g$ into $f$; order matters, since $f\circ g\ne g\circ f$ in general.
\lh{The chain of domains}An input must be valid for $g$, and $g(x)$ must in turn be valid for $f$, so the composite's domain can be restricted.
\lh{Building and undoing}Composition builds complex functions from simple ones, and inverse functions are precisely those that compose to give $x$.
\end{learnbox}

\sech{Worked Examples}

\begin{example}{ex0}{Example 1 --- Compose}
$f=x^2,\ g=x+1$. Find $f(g(x))$.\soln Replace $f$'s input with $g(x)$: $(x+1)^2$. This is the parabola shifted left 1:\eplot{-4}{2}{-1}{9}{\addplot[exblue,very thick,domain=-4:1.9]{(x+1)^2};}
\end{example}

\begin{example}{ex1}{Example 2 --- Reverse order}
Find $g(f(x))$.\soln $g(x^2)=x^2+1$ (different from $f(g(x))$).
\end{example}

\begin{example}{ex2}{Example 3 --- Evaluate}
Find $f(g(2))$.\soln $g(2)=3$, then $f(3)=9$.
\end{example}

\begin{example}{ex3}{Example 4 --- Decompose}
Write $h(x)=(x+1)^2$ as $f(g(x))$.\soln Inner $g(x)=x+1$, outer $f(x)=x^2$.
\end{example}

\begin{example}{ex4}{Example 5 --- Domain of a composite}
$f=\sqrt x,\ g=x-3$. Find $f(g(x))$ and its domain.\soln $\sqrt{x-3}$; need $x-3\ge0$, so $x\ge3$.
\end{example}

\begin{example}{ex5}{Example 6 --- Compose}
$f=x^2,\ g=x-2$. Find $f(g(x))$.\soln $(x-2)^2$.
\end{example}

\begin{example}{ex6}{Example 7 --- Reverse order}
Same $f,g$: find $g(f(x))$.\soln $x^2-2$.
\end{example}

\begin{example}{ex7}{Example 8 --- Evaluate}
$f=2x,\ g=x+3$. Find $f(g(1))$.\soln $g(1)=4,\ f(4)=8$.
\end{example}

\begin{example}{ex8}{Example 9 --- Decompose}
Decompose $h(x)=\sqrt{x+5}$.\soln $f(x)=\sqrt x,\ g(x)=x+5$.
\end{example}

\clearpage
\sech{Your Turn --- Show Your Work}
For each question, show all steps clearly in the space provided.

\begin{qbox}{Question 1}
$f=x^2,\ g=x-2$. Find $f(g(x))$.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 2}
Same $f,g$: find $g(f(x))$.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 3}
$f=2x,\ g=x+3$. Find $f(g(1))$.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 4}
Decompose $h(x)=\sqrt{x+5}$.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 5}
$f=\sqrt x,\ g=x-1$. Domain of $f(g(x))$?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 6}
$f=x+1,\ g=x^2$. Find $f(g(x))$.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 7}
Same $f,g$: find $g(f(x))$.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 8}
$f=3x,\ g=x-1$. Find $f(g(2))$.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 9}
Decompose $h(x)=(2x-1)^2$.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 10}
$f=x^2,\ g=x+4$. Find $f(g(0))$.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 11}
Is $f(g(x))$ usually equal to $g(f(x))$?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 12}
$f=\tfrac1x,\ g=x-2$. Find $f(g(x))$ and the restriction.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 13 --- Challenge}
$f=x^2,\ g=\sqrt x$. Find $f(g(x))$ and state its domain.\workspace{3cm}
\end{qbox}

\clearpage
\sech{Question Recap \& Answer Key}

\begin{recapbox}
\begin{multicols}{2}
\small
\begin{enumerate}[leftmargin=*,itemsep=3pt]
  \item $f=x^2,\ g=x-2$. Find $f(g(x))$.
  \item Same $f,g$: find $g(f(x))$.
  \item $f=2x,\ g=x+3$. Find $f(g(1))$.
  \item Decompose $h(x)=\sqrt{x+5}$.
  \item $f=\sqrt x,\ g=x-1$. Domain of $f(g(x))$?
  \item $f=x+1,\ g=x^2$. Find $f(g(x))$.
  \item Same $f,g$: find $g(f(x))$.
  \item $f=3x,\ g=x-1$. Find $f(g(2))$.
  \item Decompose $h(x)=(2x-1)^2$.
  \item $f=x^2,\ g=x+4$. Find $f(g(0))$.
  \item Is $f(g(x))$ usually equal to $g(f(x))$?
  \item $f=\tfrac1x,\ g=x-2$. Find $f(g(x))$ and the restriction.
  \item $f=x^2,\ g=\sqrt x$. Find $f(g(x))$ and state its domain.
\end{enumerate}
\end{multicols}
\end{recapbox}

\begin{keybox}
\begin{multicols}{2}
\small
\begin{enumerate}[leftmargin=*,itemsep=3pt]
  \item $(x-2)^2$
  \item $x^2-2$
  \item $8$
  \item $f=\sqrt x,\ g=x+5$
  \item $x\ge1$
  \item $x^2+1$
  \item $(x+1)^2$
  \item $3$
  \item $f=x^2,\ g=2x-1$
  \item $16$
  \item no
  \item $\dfrac{1}{x-2}$, $x\ne2$
  \item $x$, domain $x\ge0$
\end{enumerate}
\end{multicols}
\end{keybox}

\end{document}
