\documentclass[11pt]{article}
\input{_preamble}
\fancyhead[R]{\footnotesize MDM4U \textperiodcentered{} 1.2}
\begin{document}

\begin{center}
{\LARGE\bfseries\color{titleblue}Worksheet 1.2 --- Factorials \& Permutations of Distinct Objects}\\[2pt]
{\small Grade 12 Mathematics of Data Management (MDM4U) \textperiodcentered{} Unit 1: Counting, Permutations \& Combinations}
\end{center}

\begin{ideabox}
$n!$ counts full orderings of $n$ distinct objects; $_nP_r=\dfrac{n!}{(n-r)!}$ counts \emph{ordered} selections of $r$ from $n$. Simplify factorial fractions by cancelling the common tail.
\begin{itemize}[leftmargin=*,itemsep=2pt]
  \item $n!=n(n-1)\cdots2\cdot1$, and $0!=1$.
  \item $_nP_r=\dfrac{n!}{(n-r)!}$ for ordered selections.
  \item Solve $_nP_r=k$ by expanding into a polynomial in $n$.
\end{itemize}
\end{ideabox}

\begin{learnbox}
\lh{What a factorial counts}$n! = n(n-1)\cdots2\cdot1$ is the number of ways to arrange $n$ distinct objects in a row. We define $0!=1$ so that the permutation formula keeps working at the edges.
\lh{Permutations are ordered}A \textbf{permutation} is an arrangement where order matters. Filling $r$ ordered positions from $n$ distinct objects can be done in ${}_{n}P_{r}=\dfrac{n!}{(n-r)!}$ ways — the product of the shrinking choices $n(n-1)\cdots$ for $r$ factors.
\lh{Spot the keyword}If rearranging the same objects counts as a new result (1st, 2nd, 3rd place), you are counting permutations. The words \emph{arrange} and \emph{order} are the signal.
\end{learnbox}

\sech{Worked Examples}

\begin{example}{ex0}{Example 1 --- Evaluate and interpret}
Compute $6!$ and say what it counts.\soln $6!=720$ --- the number of orders of 6 distinct books on a shelf.
\end{example}

\begin{example}{ex1}{Example 2 --- Simplify a factorial fraction}
Simplify $\dfrac{10!}{7!}$ without a calculator.\soln Cancel the common tail $7!$: $10\cdot9\cdot8=720$.
\end{example}

\begin{example}{ex2}{Example 3 --- Ordered selection in context}
8 runners race for gold, silver, and bronze. How many podiums?\soln $_8P_3=8\cdot7\cdot6=336$.
\end{example}

\begin{example}{ex3}{Example 4 --- Order all objects}
How many arrangements of the letters of MONEY?\soln All 5 distinct: $5!=120$.
\end{example}

\begin{example}{ex4}{Example 5 --- Solve for $n$}
Solve $_nP_2=56$.\soln $n(n-1)=56\Rightarrow n^2-n-56=0\Rightarrow(n-8)(n+7)=0\Rightarrow n=8$.
\end{example}

\begin{example}{ex5}{Example 6 --- Algebraic simplification}
Simplify $\dfrac{(n+1)!}{(n-1)!}$.\soln $(n+1)\,n=n^2+n$.
\end{example}

\begin{example}{ex6}{Example 7 --- Assign distinct prizes}
5 different trophies are given to 8 students, at most one each. How many ways?\soln Ordered selection: $_8P_5=8\cdot7\cdot6\cdot5\cdot4=6720$.
\end{example}

\begin{example}{ex7}{Example 8 --- Arrangement with a fixed spot}
7 people line up for a photo with one specific person in the middle (4th) position. How many line-ups?\soln Fix that person; arrange the other 6: $6!=720$.
\end{example}

\begin{example}{ex8}{Example 9 --- Two-part count}
A gallery hangs 4 of 9 distinct paintings in a row. \\ \textbf{(a)} How many displays? \\ \textbf{(b)} How many put a specific painting first?\soln (a) $_9P_4=3024$. (b) fix it first; arrange 3 of the other 8: $_8P_3=336$.
\end{example}

\clearpage
\sech{Your Turn --- Show Your Work}
For each question, show all steps clearly in the space provided.

\begin{qbox}{Question 1}
Evaluate $8!$.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 2}
In how many orders can 7 distinct parade floats line up?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 3}
Simplify $\dfrac{11!}{8!}$.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 4}
A club elects a president, then a VP, then a treasurer from 12 members. How many outcomes?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 5}
Find $_{10}P_3$.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 6}
Solve $_nP_2=42$.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 7}
How many arrangements of the letters of PLANET (all distinct)?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 8}
A 4-character password uses distinct letters from $\{A,B,C,D,E,F\}$. How many?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 9}
Simplify $\dfrac{n!}{(n-3)!}$.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 10}
Solve $_nP_3=210$.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 11}
In how many ways can gold and silver medals be given among 15 athletes?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 12}
5 distinct books sit on a shelf. \\ \textbf{(a)} How many arrangements? \\ \textbf{(b)} How many with a particular book at an end?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 13 --- Challenge}
Simplify $\dfrac{(n+2)!}{n!}$, then solve $\dfrac{(n+2)!}{n!}=72$.\workspace{3cm}
\end{qbox}

\clearpage
\sech{Question Recap \& Answer Key}

\begin{recapbox}
\begin{multicols}{2}
\small
\begin{enumerate}[leftmargin=*,itemsep=3pt]
  \item Evaluate $8!$.
  \item In how many orders can 7 distinct parade floats line up?
  \item Simplify $\dfrac{11!}{8!}$.
  \item A club elects a president, then a VP, then a treasurer from 12 members. How many outcomes?
  \item Find $_{10}P_3$.
  \item Solve $_nP_2=42$.
  \item How many arrangements of the letters of PLANET (all distinct)?
  \item A 4-character password uses distinct letters from $\{A,B,C,D,E,F\}$. How many?
  \item Simplify $\dfrac{n!}{(n-3)!}$.
  \item Solve $_nP_3=210$.
  \item In how many ways can gold and silver medals be given among 15 athletes?
  \item 5 distinct books sit on a shelf. \\ \textbf{(a)} How many arrangements? \\ \textbf{(b)} How many with a particular book at an end?
  \item Simplify $\dfrac{(n+2)!}{n!}$, then solve $\dfrac{(n+2)!}{n!}=72$.
\end{enumerate}
\end{multicols}
\end{recapbox}

\begin{keybox}
\begin{multicols}{2}
\small
\begin{enumerate}[leftmargin=*,itemsep=3pt]
  \item $40\,320$
  \item $7!=5040$
  \item $11\cdot10\cdot9=990$
  \item $12\cdot11\cdot10=1320$
  \item $720$
  \item $n=7$
  \item $720$
  \item $_6P_4=360$
  \item $n(n-1)(n-2)$
  \item $n=7$
  \item $_{15}P_2=210$
  \item (a) $120$; (b) $2\cdot4!=48$
  \item $(n+2)(n+1)$; $=72\Rightarrow n=7$
\end{enumerate}
\end{multicols}
\end{keybox}

\end{document}
