\documentclass[11pt]{article}
\input{_preamble}
\fancyhead[R]{\footnotesize MDM4U \textperiodcentered{} 6.5}
\begin{document}

\begin{center}
{\LARGE\bfseries\color{titleblue}Worksheet 6.5 --- Correlation vs Causation}\\[2pt]
{\small Grade 12 Mathematics of Data Management (MDM4U) \textperiodcentered{} Unit 6: Two-Variable Statistics}
\end{center}

\begin{ideabox}
A strong correlation never, by itself, proves cause and effect. A hidden \emph{confounding} (lurking) variable, reverse causation, or plain coincidence can all create correlation. Only a controlled experiment can establish causation.
\begin{itemize}[leftmargin=*,itemsep=2pt]
  \item Correlation $\ne$ causation.
  \item A confounding (lurking) variable can drive both quantities.
  \item Also possible: reverse causation, common cause, or coincidence.
\end{itemize}
\end{ideabox}

\begin{learnbox}
\lh{A crucial distinction}A strong correlation shows two variables moving together; it does \textbf{not} prove that one causes the other.
\lh{Hidden variables}A \textbf{confounding} variable can drive both quantities: ice-cream sales and drownings rise together because of a shared cause, summer heat.
\lh{Establishing cause}Only a well-designed \textbf{experiment} with controls — not observation alone — can support a claim of causation.
\end{learnbox}

\sech{Worked Examples}

\begin{example}{ex0}{Example 1 --- The classic confounder}
Ice-cream sales and drownings rise together. Does ice cream cause drownings?\soln No --- hot weather (a confounder) drives both.
\end{example}

\begin{example}{ex1}{Example 2 --- Reverse causation}
Sick people take more medicine, yet medicine is meant to help. What is happening?\soln The direction is reversed --- illness drives medicine use, not the other way.
\end{example}

\begin{example}{ex2}{Example 3 --- Coincidence}
Two unrelated trends happen to move together for a few years. Explain.\soln A coincidental correlation --- no real link at all.
\end{example}

\begin{example}{ex3}{Example 4 --- Common cause}
Larger vocabulary correlates with bigger shoe size in children. Why?\soln Both grow with \emph{age} --- a common cause.
\end{example}

\begin{example}{ex4}{Example 5 --- Why experiment}
Why is a controlled experiment needed to prove causation?\soln It randomizes and controls other variables, isolating the effect of one factor.
\end{example}

\begin{example}{ex5}{Example 6 --- Spot the lurking variable}
Towns with more churches have more bars. What is the lurking variable?\soln Population size --- bigger towns have more of both.
\end{example}

\begin{example}{ex6}{Example 7 --- Observational limits}
An observational study finds a correlation. What can it conclude?\soln Correlation only --- not causation.
\end{example}

\begin{example}{ex7}{Example 8 --- Strength is not proof}
Two variables have $r=0.99$. Does that prove one causes the other?\soln No --- even a near-perfect correlation can be confounded or coincidental.
\end{example}

\begin{example}{ex8}{Example 9 --- Full reasoning (multi-part)}
Students with more books at home score higher. \\ \textbf{(a)} Does this prove books raise scores? \\ \textbf{(b)} Name a lurking variable. \\ \textbf{(c)} How could causation be tested?\soln (a) No. (b) family income/education. (c) a controlled study that varies book access while holding other factors fixed.
\end{example}

\clearpage
\sech{Your Turn --- Show Your Work}
For each question, show all steps clearly in the space provided.

\begin{qbox}{Question 1}
Does a high correlation prove causation?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 2}
Ice-cream sales and drownings both rise. Name the lurking variable.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 3}
What kind of link is \emph{a rooster crows, then the sun rises}?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 4}
What is required to establish causation?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 5}
What is a confounding variable?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 6}
Bigger shoe size correlates with better reading in children. Lurking variable?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 7}
What can an observational study establish?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 8}
Coffee drinkers live longer. Does coffee cause longevity?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 9}
Two variables both driven by a third factor illustrate what?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 10}
A randomized experiment controls for what?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 11}
A correlation of $0.99$ between two variables --- does it prove cause?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 12}
Name one alternative to causation that can explain a correlation.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 13 --- Challenge}
Cities with more police have more crime. \\ \textbf{(a)} Does policing cause crime? \\ \textbf{(b)} Name a lurking variable. \\ \textbf{(c)} How could you test the claim?\workspace{3cm}
\end{qbox}

\clearpage
\sech{Question Recap \& Answer Key}

\begin{recapbox}
\begin{multicols}{2}
\small
\begin{enumerate}[leftmargin=*,itemsep=3pt]
  \item Does a high correlation prove causation?
  \item Ice-cream sales and drownings both rise. Name the lurking variable.
  \item What kind of link is \emph{a rooster crows, then the sun rises}?
  \item What is required to establish causation?
  \item What is a confounding variable?
  \item Bigger shoe size correlates with better reading in children. Lurking variable?
  \item What can an observational study establish?
  \item Coffee drinkers live longer. Does coffee cause longevity?
  \item Two variables both driven by a third factor illustrate what?
  \item A randomized experiment controls for what?
  \item A correlation of $0.99$ between two variables --- does it prove cause?
  \item Name one alternative to causation that can explain a correlation.
  \item Cities with more police have more crime. \\ \textbf{(a)} Does policing cause crime? \\ \textbf{(b)} Name a lurking variable. \\ \textbf{(c)} How could you test the claim?
\end{enumerate}
\end{multicols}
\end{recapbox}

\begin{keybox}
\begin{multicols}{2}
\small
\begin{enumerate}[leftmargin=*,itemsep=3pt]
  \item no
  \item hot weather
  \item coincidental (and reverse) --- not causal
  \item a controlled experiment
  \item one that affects both variables
  \item age
  \item correlation, not causation
  \item not necessarily --- it may be confounded
  \item a common cause
  \item confounding variables
  \item no
  \item confounding, reverse causation, or coincidence
  \item (a) no; (b) city/population size; (c) a controlled comparison
\end{enumerate}
\end{multicols}
\end{keybox}

\end{document}
