\documentclass[11pt]{article}
\input{_preamble}
\fancyhead[R]{\footnotesize MDM4U \textperiodcentered{} 7.1}
\begin{document}

\begin{center}
{\LARGE\bfseries\color{titleblue}Worksheet 7.1 --- Designing the Study}\\[2pt]
{\small Grade 12 Mathematics of Data Management (MDM4U) \textperiodcentered{} Unit 7: The Culminating Investigation}
\end{center}

\begin{ideabox}
A good investigation turns curiosity into a \emph{researchable question} with clearly defined variables and a hypothesis, then chooses a design --- observational or experimental --- that can actually answer it.
\begin{itemize}[leftmargin=*,itemsep=2pt]
  \item A researchable question is focused, measurable, and answerable with data.
  \item Identify the explanatory and response variables and state a hypothesis.
  \item Observational study (measures as-is) vs. experiment (imposes a treatment); only an experiment can show causation.
\end{itemize}
\end{ideabox}

\begin{learnbox}
\lh{Start with a question}An investigation opens with a clear, answerable \textbf{question} and a hypothesis about what you expect to find.
\lh{Plan the data}Decide which variables to measure, which population to study, and how to sample it without bias — thoughtful design prevents useless data.
\lh{Anticipate the analysis}Choose your tools (graphs, summary statistics, regression) \emph{before} collecting, so the data you gather actually answers the question.
\end{learnbox}

\sech{Worked Examples}

\begin{example}{ex0}{Example 1 --- Refine a vague question}
Turn the vague question \emph{Is exercise good?} into a researchable one.\soln E.g. \emph{Do students who exercise at least 3 hours a week have higher fitness scores?} --- specific, measurable variables.
\end{example}

\begin{example}{ex1}{Example 2 --- Identify the variables}
In \emph{does study time affect test scores}, name the variables.\soln Explanatory: study time; response: test score.
\end{example}

\begin{example}{ex2}{Example 3 --- State a hypothesis}
Write a hypothesis for a sleep-and-grades study.\soln \emph{More sleep is associated with higher grades} --- a testable prediction of the relationship.
\end{example}

\begin{example}{ex3}{Example 4 --- Observational vs. experimental}
To test whether a fertilizer increases yield, which design proves cause?\soln An experiment: randomly assign plots to fertilizer or none and compare --- only this controls confounders.
\end{example}

\begin{example}{ex4}{Example 5 --- Population and frame}
Distinguish the target population from the sampling frame.\soln The population is everyone of interest; the frame is the list you can actually sample from (they may differ, causing coverage bias).
\end{example}

\begin{example}{ex5}{Example 6 --- Control a confounder}
A study links coffee to poor sleep, but coffee drinkers also work late. What must the design control?\soln The confounder --- work hours --- e.g. by matching or randomizing.
\end{example}

\begin{example}{ex6}{Example 7 --- One- vs. two-variable question}
Classify: (a) \emph{what is the average commute time?}; (b) \emph{does commute time depend on distance?}\soln (a) one-variable (a single measured quantity); (b) two-variable (a relationship).
\end{example}

\begin{example}{ex7}{Example 8 --- Ethics in design}
Name one ethical requirement when studying people.\soln Informed consent (also privacy/anonymity and the right to withdraw).
\end{example}

\begin{example}{ex8}{Example 9 --- Design a study (multi-part)}
For \emph{does sleep affect grades}, state \\ \textbf{(a)} the variables; \\ \textbf{(b)} a hypothesis; \\ \textbf{(c)} observational or experimental?\soln (a) sleep (explanatory), grades (response). (b) more sleep $\to$ higher grades. (c) observational (you cannot ethically force sleep amounts).
\end{example}

\clearpage
\sech{Your Turn --- Show Your Work}
For each question, show all steps clearly in the space provided.

\begin{qbox}{Question 1}
Rewrite the vague question \emph{are phones bad?} as a researchable one.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 2}
Name the explanatory and response variables in \emph{does temperature affect ice-cream sales?}\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 3}
Write a testable hypothesis about practice time and free-throw success.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 4}
Which design can establish causation: observational or experimental?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 5}
Classify \emph{what is the median household income?} --- one- or two-variable?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 6}
Classify \emph{is height related to shoe size?} --- one- or two-variable?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 7}
A study links screen time to grades, but both relate to age. Name the confounder.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 8}
State one ethical requirement when collecting data from people.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 9}
What is the difference between a population and a sampling frame?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 10}
To prove a drug works, why randomly assign patients to drug vs. placebo?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 11}
Give the response variable in \emph{does fertilizer amount change plant height?}\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 12}
Is \emph{how many hours do teens sleep?} a one- or two-variable question?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 13 --- Challenge}
For \emph{does music while studying affect recall}, state \\ \textbf{(a)} the variables; \\ \textbf{(b)} a hypothesis; \\ \textbf{(c)} a design that could show causation.\workspace{3cm}
\end{qbox}

\clearpage
\sech{Question Recap \& Answer Key}

\begin{recapbox}
\begin{multicols}{2}
\small
\begin{enumerate}[leftmargin=*,itemsep=3pt]
  \item Rewrite the vague question \emph{are phones bad?} as a researchable one.
  \item Name the explanatory and response variables in \emph{does temperature affect ice-cream sales?}
  \item Write a testable hypothesis about practice time and free-throw success.
  \item Which design can establish causation: observational or experimental?
  \item Classify \emph{what is the median household income?} --- one- or two-variable?
  \item Classify \emph{is height related to shoe size?} --- one- or two-variable?
  \item A study links screen time to grades, but both relate to age. Name the confounder.
  \item State one ethical requirement when collecting data from people.
  \item What is the difference between a population and a sampling frame?
  \item To prove a drug works, why randomly assign patients to drug vs. placebo?
  \item Give the response variable in \emph{does fertilizer amount change plant height?}
  \item Is \emph{how many hours do teens sleep?} a one- or two-variable question?
  \item For \emph{does music while studying affect recall}, state \\ \textbf{(a)} the variables; \\ \textbf{(b)} a hypothesis; \\ \textbf{(c)} a design that could show causation.
\end{enumerate}
\end{multicols}
\end{recapbox}

\begin{keybox}
\begin{multicols}{2}
\small
\begin{enumerate}[leftmargin=*,itemsep=3pt]
  \item e.g. do teens using phones $>4$ h/day sleep fewer hours?
  \item explanatory: temperature; response: sales
  \item e.g. more practice raises free-throw percentage
  \item experimental
  \item one-variable
  \item two-variable
  \item age
  \item informed consent (or privacy/anonymity)
  \item population $=$ all of interest; frame $=$ the list you can sample
  \item to control confounders so the groups differ only by treatment
  \item plant height
  \item one-variable
  \item (a) music (expl.), recall (resp.); (b) music lowers recall; (c) a randomized experiment
\end{enumerate}
\end{multicols}
\end{keybox}

\end{document}
