\documentclass[11pt]{article}
\input{_preamble}
\fancyhead[R]{\footnotesize MDM4U \textperiodcentered{} 7.2}
\begin{document}

\begin{center}
{\LARGE\bfseries\color{titleblue}Worksheet 7.2 --- Collecting \& Managing Data}\\[2pt]
{\small Grade 12 Mathematics of Data Management (MDM4U) \textperiodcentered{} Unit 7: The Culminating Investigation}
\end{center}

\begin{ideabox}
Data may be \emph{primary} (you collect it) or \emph{secondary} (already published). Before analysis it must be cleaned --- fixing errors, outliers, and missing values --- and handled ethically, with sources cited.
\begin{itemize}[leftmargin=*,itemsep=2pt]
  \item Primary (first-hand) vs. secondary (existing) data.
  \item Data types: categorical vs. numerical; discrete vs. continuous.
  \item Clean errors/outliers/missing values; respect privacy and cite sources.
\end{itemize}
\end{ideabox}

\begin{learnbox}
\lh{Gathering data}Collect data through surveys, experiments, or existing databases, following your sampling plan so the sample stays representative.
\lh{Organizing it}Record everything in a clean, structured table that can be sorted, filtered, and analyzed; tidy organization prevents errors down the line.
\lh{Checking quality}Scan for missing entries, outliers, and measurement mistakes — deciding how to handle them honestly is part of managing data well.
\end{learnbox}

\sech{Worked Examples}

\begin{example}{ex0}{Example 1 --- Primary vs. secondary}
You survey classmates about sleep. Primary or secondary?\soln Primary --- you collected it first-hand.
\end{example}

\begin{example}{ex1}{Example 2 --- Using published data}
You download census tables. Primary or secondary?\soln Secondary --- it was collected by someone else.
\end{example}

\begin{example}{ex2}{Example 3 --- Spot a data error}
A list of ages includes $200$. What should you do?\soln Flag it as a likely entry error and correct or remove it before analysis.
\end{example}

\begin{example}{ex3}{Example 4 --- Missing values}
Some survey rows have blank income. Name two options.\soln Exclude those rows, or impute a reasonable value --- and report what you did.
\end{example}

\begin{example}{ex4}{Example 5 --- Classify data types}
Classify: (a) eye colour; (b) number of siblings; (c) height.\soln (a) categorical; (b) numerical discrete; (c) numerical continuous.
\end{example}

\begin{example}{ex5}{Example 6 --- Ethics: privacy}
Why anonymize survey responses?\soln To protect participants' privacy and encourage honest answers.
\end{example}

\begin{example}{ex6}{Example 7 --- Cite a source}
Why must a secondary source be cited?\soln To give credit, allow verification, and let readers judge the data's reliability.
\end{example}

\begin{example}{ex7}{Example 8 --- Consistent units}
Half the heights are in cm, half in inches. What is the risk?\soln Mixed units corrupt any calculation; convert everything to one unit first.
\end{example}

\begin{example}{ex8}{Example 9 --- Clean a dataset (multi-part)}
A downloaded table lists a person's age as $200$. \\ \textbf{(a)} primary or secondary? \\ \textbf{(b)} what is the issue? \\ \textbf{(c)} how do you handle it?\soln (a) secondary. (b) an impossible outlier (error). (c) verify, then correct or remove it and note the change.
\end{example}

\clearpage
\sech{Your Turn --- Show Your Work}
For each question, show all steps clearly in the space provided.

\begin{qbox}{Question 1}
You measure plant heights yourself. Primary or secondary data?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 2}
You use Statistics Canada tables. Primary or secondary?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 3}
Classify: favourite colour --- categorical or numerical?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 4}
Classify: number of pets --- discrete or continuous?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 5}
Classify: exact weight of a package --- discrete or continuous?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 6}
A test score of $150$ appears on a $100$-point exam. What is it?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 7}
Name one way to handle missing values in a dataset.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 8}
Why should survey responses be anonymized?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 9}
Why cite a secondary data source?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 10}
Heights are recorded in both cm and inches. What must you do first?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 11}
Classify: brand of phone owned --- categorical or numerical?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 12}
Give one example of primary data collection.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 13 --- Challenge}
A table lists an age of $200$. \\ \textbf{(a)} primary or secondary? \\ \textbf{(b)} the issue? \\ \textbf{(c)} how to handle it?\workspace{3cm}
\end{qbox}

\clearpage
\sech{Question Recap \& Answer Key}

\begin{recapbox}
\begin{multicols}{2}
\small
\begin{enumerate}[leftmargin=*,itemsep=3pt]
  \item You measure plant heights yourself. Primary or secondary data?
  \item You use Statistics Canada tables. Primary or secondary?
  \item Classify: favourite colour --- categorical or numerical?
  \item Classify: number of pets --- discrete or continuous?
  \item Classify: exact weight of a package --- discrete or continuous?
  \item A test score of $150$ appears on a $100$-point exam. What is it?
  \item Name one way to handle missing values in a dataset.
  \item Why should survey responses be anonymized?
  \item Why cite a secondary data source?
  \item Heights are recorded in both cm and inches. What must you do first?
  \item Classify: brand of phone owned --- categorical or numerical?
  \item Give one example of primary data collection.
  \item A table lists an age of $200$. \\ \textbf{(a)} primary or secondary? \\ \textbf{(b)} the issue? \\ \textbf{(c)} how to handle it?
\end{enumerate}
\end{multicols}
\end{recapbox}

\begin{keybox}
\begin{multicols}{2}
\small
\begin{enumerate}[leftmargin=*,itemsep=3pt]
  \item primary
  \item secondary
  \item categorical
  \item discrete
  \item continuous
  \item an error/outlier (impossible value)
  \item exclude the rows, or impute and report it
  \item to protect privacy and encourage honesty
  \item credit, verification, and reliability
  \item convert all to a single unit
  \item categorical
  \item e.g. running your own survey/experiment
  \item (a) secondary; (b) impossible outlier; (c) verify then correct/remove
\end{enumerate}
\end{multicols}
\end{keybox}

\end{document}
