\documentclass[11pt]{article}
\input{_preamble}
\fancyhead[R]{\footnotesize MDM4U \textperiodcentered{} 7.3}
\begin{document}

\begin{center}
{\LARGE\bfseries\color{titleblue}Worksheet 7.3 --- Analyzing the Data}\\[2pt]
{\small Grade 12 Mathematics of Data Management (MDM4U) \textperiodcentered{} Unit 7: The Culminating Investigation}
\end{center}

\begin{ideabox}
Analysis puts the whole course to work: choose the right one-variable summary (centre, spread, shape) and two-variable tool ($r$, regression), while staying alert to bias and variability.
\begin{itemize}[leftmargin=*,itemsep=2pt]
  \item Pick the centre/spread that suit the shape (median \& IQR for skewed data).
  \item Two-variable: use $r$ for strength and a regression line to predict.
  \item Interpret results with an eye on sampling bias and variability.
\end{itemize}
\end{ideabox}

\begin{learnbox}
\lh{Summarize first}Begin with \textbf{descriptive} tools — measures of centre and spread, plus graphs — to see the overall picture before digging deeper.
\lh{Look for relationships}Use scatter plots, correlation, and regression to test whether variables are linked, and probability where chance plays a role.
\lh{Interpret in context}Statistics only matter when tied back to the original question — always translate the numbers into a plain-language finding.
\end{learnbox}

\sech{Worked Examples}

\begin{example}{ex0}{Example 1 --- Choose the centre}
Incomes are strongly right-skewed. Which centre?\soln The median --- it resists the high outliers.
\end{example}

\begin{example}{ex1}{Example 2 --- Choose the spread}
For skewed data with outliers, which spread?\soln The IQR --- it is resistant, unlike the range or $\sigma$.
\end{example}

\begin{example}{ex2}{Example 3 --- Summarize one variable}
Find the median of $\{2,3,3,4,50\}$.\soln $3$ --- and note the $50$ is an outlier inflating the mean.
\end{example}

\begin{example}{ex3}{Example 4 --- Two-variable strength}
Technology reports $r=0.9$ for study time vs. score. Interpret.\soln A strong positive linear relationship.
\end{example}

\begin{example}{ex4}{Example 5 --- Predict with regression}
Using $\hat y=2x+5$, predict the score at $x=10$ hours.\soln $\hat y=25$.
\end{example}

\begin{example}{ex5}{Example 6 --- Compare two groups}
How can you compare two classes' test results?\soln Compare centres and spreads --- e.g. side-by-side box plots or means with $\sigma$.
\end{example}

\begin{example}{ex6}{Example 7 --- Interpret r squared}
A model has $r^{2}=0.81$. What does it say?\soln $81\%$ of the variation in the response is explained by the model.
\end{example}

\begin{example}{ex7}{Example 8 --- Bias check}
Your data came from a volunteer web poll. How does that affect analysis?\soln Self-selection bias --- results may not generalize to the whole population.
\end{example}

\begin{example}{ex8}{Example 9 --- Full analysis (multi-part)}
For $\{4,6,6,8,11\}$, find \\ \textbf{(a)} the mean; \\ \textbf{(b)} the median; \\ \textbf{(c)} which better describes the centre.\soln (a) $7$. (b) $6$. (c) both are close --- the mean is fine (little skew).
\end{example}

\clearpage
\sech{Your Turn --- Show Your Work}
For each question, show all steps clearly in the space provided.

\begin{qbox}{Question 1}
For strongly skewed data, which centre is best?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 2}
For data with outliers, which spread is most resistant?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 3}
Find the median of $\{2,3,3,4,50\}$.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 4}
Technology gives $r=-0.8$. Describe the relationship.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 5}
Using $\hat y=3x+2$, predict $y$ at $x=5$.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 6}
A model has $r^{2}=0.64$. What percent of variation is explained?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 7}
Name a graph for comparing two groups' distributions.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 8}
Data came from a volunteer online poll. Name the bias.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 9}
Find the mean of $\{4,6,6,8,11\}$.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 10}
Would a scatter plot or a histogram show a relationship between two variables?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 11}
$r=0.2$: is the linear relationship strong or weak?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 12}
For symmetric data with no outliers, which centre is fine to use?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 13 --- Challenge}
For $\{4,6,6,8,11\}$, find \\ \textbf{(a)} the mean; \\ \textbf{(b)} the median; \\ \textbf{(c)} which better describes the centre.\workspace{3cm}
\end{qbox}

\clearpage
\sech{Question Recap \& Answer Key}

\begin{recapbox}
\begin{multicols}{2}
\small
\begin{enumerate}[leftmargin=*,itemsep=3pt]
  \item For strongly skewed data, which centre is best?
  \item For data with outliers, which spread is most resistant?
  \item Find the median of $\{2,3,3,4,50\}$.
  \item Technology gives $r=-0.8$. Describe the relationship.
  \item Using $\hat y=3x+2$, predict $y$ at $x=5$.
  \item A model has $r^{2}=0.64$. What percent of variation is explained?
  \item Name a graph for comparing two groups' distributions.
  \item Data came from a volunteer online poll. Name the bias.
  \item Find the mean of $\{4,6,6,8,11\}$.
  \item Would a scatter plot or a histogram show a relationship between two variables?
  \item $r=0.2$: is the linear relationship strong or weak?
  \item For symmetric data with no outliers, which centre is fine to use?
  \item For $\{4,6,6,8,11\}$, find \\ \textbf{(a)} the mean; \\ \textbf{(b)} the median; \\ \textbf{(c)} which better describes the centre.
\end{enumerate}
\end{multicols}
\end{recapbox}

\begin{keybox}
\begin{multicols}{2}
\small
\begin{enumerate}[leftmargin=*,itemsep=3pt]
  \item the median
  \item the IQR
  \item $3$
  \item strong negative linear
  \item $17$
  \item $64\%$
  \item side-by-side box plots
  \item self-selection (voluntary-response) bias
  \item $\tfrac{35}{5}=7$
  \item a scatter plot
  \item weak
  \item the mean
  \item (a) $7$; (b) $6$; (c) either --- little skew
\end{enumerate}
\end{multicols}
\end{keybox}

\end{document}
