\documentclass[11pt]{article}
\input{_preamble}
\fancyhead[R]{\footnotesize MDM4U \textperiodcentered{} 7.4}
\begin{document}

\begin{center}
{\LARGE\bfseries\color{titleblue}Worksheet 7.4 --- Communicating Conclusions}\\[2pt]
{\small Grade 12 Mathematics of Data Management (MDM4U) \textperiodcentered{} Unit 7: The Culminating Investigation}
\end{center}

\begin{ideabox}
A result is only as strong as its argument: interpret in real-world context, acknowledge the limitations and assumptions, and never claim more than the data support (correlation is not causation).
\begin{itemize}[leftmargin=*,itemsep=2pt]
  \item Interpret findings in context, with appropriate units and scale.
  \item State limitations: sample size, bias, variability, and the danger of extrapolation.
  \item Match the claim to the evidence --- an observational study shows correlation, not cause.
\end{itemize}
\end{ideabox}

\begin{learnbox}
\lh{Answer the question}A conclusion should directly answer the original question, backed by your evidence — not merely restate the numbers you found.
\lh{Show the evidence}Support the argument with well-labelled graphs and the key statistics, so a reader can see why the conclusion follows.
\lh{Name the limitations}Strong reports state their \textbf{limitations} — sample size, possible bias, the danger of extrapolation — and avoid over-claiming causation.
\end{learnbox}

\sech{Worked Examples}

\begin{example}{ex0}{Example 1 --- Interpret in context}
A model gives a slope of $2$ marks per study-hour. State it in plain language.\soln Each additional hour of study is associated with about $2$ more marks.
\end{example}

\begin{example}{ex1}{Example 2 --- Acknowledge a limitation}
A survey of 15 friends finds a trend. What limitation must you state?\soln The tiny, non-random sample --- results may not generalize.
\end{example}

\begin{example}{ex2}{Example 3 --- Avoid a causation overclaim}
Data show a correlation between two variables. What may you \emph{not} conclude?\soln That one causes the other --- correlation alone does not prove causation.
\end{example}

\begin{example}{ex3}{Example 4 --- Note extrapolation}
Your data cover ages $10$--$18$. Why not predict for age $40$?\soln That is extrapolation far outside the data; the pattern may not hold.
\end{example}

\begin{example}{ex4}{Example 5 --- Choose the right graph}
You want to show one variable's shape. Which display?\soln A histogram or box plot --- not a scatter plot (that is for relationships).
\end{example}

\begin{example}{ex5}{Example 6 --- Report uncertainty}
Why report a margin of error with a poll result?\soln It conveys the sampling variability --- the true value likely lies within that range.
\end{example}

\begin{example}{ex6}{Example 7 --- Address bias in the conclusion}
Your sample over-represented one group. How should the conclusion reflect this?\soln State that the bias may skew results and limit how far they generalize.
\end{example}

\begin{example}{ex7}{Example 8 --- Practical vs. statistical}
A huge study finds a real but tiny effect. What should you note?\soln It may be statistically detectable yet too small to matter practically.
\end{example}

\begin{example}{ex8}{Example 9 --- Critique a conclusion (multi-part)}
A survey concludes \emph{video games cause better reflexes}. \\ \textbf{(a)} what is wrong? \\ \textbf{(b)} a limitation? \\ \textbf{(c)} reword it responsibly.\soln (a) it claims causation from a survey (correlation only). (b) possible confounders/self-report. (c) \emph{gaming is associated with faster reflexes in this sample}.
\end{example}

\clearpage
\sech{Your Turn --- Show Your Work}
For each question, show all steps clearly in the space provided.

\begin{qbox}{Question 1}
A regression slope is $5$ dollars per item. State its meaning in words.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 2}
A study surveys only 12 people. What limitation must be reported?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 3}
Two variables are correlated. What conclusion is NOT justified?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 4}
Data span years $1$--$5$. Why distrust a prediction for year $20$?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 5}
Which display shows a single variable's shape: histogram or scatter plot?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 6}
Why report a margin of error with a poll?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 7}
A sample over-represents one region. How should the conclusion note this?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 8}
A large study finds a tiny effect. What distinction should you make?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 9}
Rewrite the claim \emph{sunscreen causes more beach trips} to avoid a false causal statement.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 10}
Name one assumption worth stating in a conclusion.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 11}
Which is appropriate for an observational study: \emph{linked to} or \emph{causes}?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 12}
Why include units when reporting a mean?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 13 --- Challenge}
A survey concludes \emph{phones cause lower grades}. \\ \textbf{(a)} what is wrong? \\ \textbf{(b)} a limitation? \\ \textbf{(c)} reword it responsibly.\workspace{3cm}
\end{qbox}

\clearpage
\sech{Question Recap \& Answer Key}

\begin{recapbox}
\begin{multicols}{2}
\small
\begin{enumerate}[leftmargin=*,itemsep=3pt]
  \item A regression slope is $5$ dollars per item. State its meaning in words.
  \item A study surveys only 12 people. What limitation must be reported?
  \item Two variables are correlated. What conclusion is NOT justified?
  \item Data span years $1$--$5$. Why distrust a prediction for year $20$?
  \item Which display shows a single variable's shape: histogram or scatter plot?
  \item Why report a margin of error with a poll?
  \item A sample over-represents one region. How should the conclusion note this?
  \item A large study finds a tiny effect. What distinction should you make?
  \item Rewrite the claim \emph{sunscreen causes more beach trips} to avoid a false causal statement.
  \item Name one assumption worth stating in a conclusion.
  \item Which is appropriate for an observational study: \emph{linked to} or \emph{causes}?
  \item Why include units when reporting a mean?
  \item A survey concludes \emph{phones cause lower grades}. \\ \textbf{(a)} what is wrong? \\ \textbf{(b)} a limitation? \\ \textbf{(c)} reword it responsibly.
\end{enumerate}
\end{multicols}
\end{recapbox}

\begin{keybox}
\begin{multicols}{2}
\small
\begin{enumerate}[leftmargin=*,itemsep=3pt]
  \item each extra item adds \$5 on average
  \item the small, likely non-random sample
  \item that one variable causes the other
  \item it is extrapolation far outside the data
  \item histogram
  \item it shows the sampling variability/uncertainty
  \item state that the bias limits how far it generalizes
  \item statistical vs. practical significance
  \item e.g. sunscreen use is associated with more beach trips
  \item e.g. the sample is representative, or units are consistent
  \item linked to
  \item to give the numbers meaning and scale
  \item (a) causation claimed from a survey; (b) confounders/self-report; (c) phone use is associated with lower grades here
\end{enumerate}
\end{multicols}
\end{keybox}

\end{document}
