\documentclass[11pt]{article}
\input{_preamble}
\fancyhead[R]{\footnotesize MAP4C \textperiodcentered{} 1.5}
\begin{document}

\begin{center}
{\LARGE\bfseries\color{titleblue}Worksheet 1.5 --- Interpreting Graphs \& Describing Trends}\\[2pt]
{\small Grade 12 Foundations for College Mathematics (MAP4C) \textperiodcentered{} Unit 1: Mathematical Models}
\end{center}

\begin{ideabox}
Read a graph to describe how two quantities relate, identify the trend, and make predictions — always with correct units and stated assumptions.
\begin{itemize}[leftmargin=*,itemsep=2pt]
  \item Describe direction: rise, fall, level, reverse.
  \item Trend $=$ overall direction; extrapolate to predict.
  \item Note maxima, minima, and intercepts.
\end{itemize}
\end{ideabox}

\begin{learnbox}
\lh{Describe the relationship}Say how the output changes as the input rises — increasing, decreasing, or levelling off — with the correct quantities and units.
\lh{Trends and predictions}Extend a trend to predict, and always state the assumption that it continues.
\lh{Special points}Note maxima, minima, and intercepts, and be cautious extrapolating a curved trend far ahead.
\end{learnbox}

\sech{Worked Examples}

\begin{example}{ex0}{Example 1 --- Distance–time}
A distance–time graph is a straight line from $(0,0)$ to $(2,120)$ (km, h). Describe it.\soln Constant speed; rate $=\tfrac{120}{2}=60$ km/h.
\end{example}

\begin{example}{ex1}{Example 2 --- Levelling off}
A cooling graph falls fast then flattens. Describe the trend.\soln Decreasing, then levelling toward room temperature (exponential decay to a limit).
\end{example}

\begin{example}{ex2}{Example 3 --- Linear prediction}
Sales (\$000) are $20,24,28,32$ for years $1$–$4$. Predict year $6$.\soln Rises \$4000/yr (linear): year $5\to36$, year $6\to40$; assuming the trend continues.
\end{example}

\begin{example}{ex3}{Example 4 --- Reading a maximum}
A profit graph peaks at $(50,900)$. Interpret the peak.\soln Maximum profit \$900 at $50$ units; profit falls beyond that.
\end{example}

\begin{example}{ex4}{Example 5 --- Exponential trend}
Users are $10,20,40,80$ over four months. Predict month $5$.\soln Doubling (exponential): $160$.
\end{example}

\begin{example}{ex5}{Example 6 --- Rate direction}
A temperature graph falls and flattens. Is the size of the rate increasing or decreasing?\soln Decreasing — the drop slows.
\end{example}

\begin{example}{ex6}{Example 7 --- Cautious extrapolation}
A world-record-time graph trends down but flattens. Why is a $2100$ prediction risky?\soln Improvements shrink; extending far ahead may give impossible times.
\end{example}

\begin{example}{ex7}{Example 8 --- Intercept meaning}
A cost graph crosses the vertical axis at \$20. What does that represent?\soln A fixed starting cost (base fee) when usage is $0$.
\end{example}

\begin{example}{ex8}{Example 9 --- Compare two graphs}
Graph A is a straight line; graph B curves upward. Compare their trends.\soln A: constant rate; B: increasing rate (accelerating).
\end{example}

\clearpage
\sech{Your Turn --- Show Your Work}
For each question, show all steps clearly in the space provided.

\begin{qbox}{Question 1}
A graph rises then stays flat. Describe the trend.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 2}
A distance–time line goes $(0,0)$ to $(4,200)$. Find the speed with units.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 3}
Sales are $10,20,40,80$ over four years. Predict year $5$ and name the pattern.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 4}
A graph falls and flattens. Is the size of the rate of change increasing or decreasing?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 5}
A profit graph peaks at $(30,450)$. What does the peak mean?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 6}
Sales rise by \$5000 each year from \$25\,000. Predict after $3$ more years.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 7}
Why should a prediction far beyond the data be treated with caution?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 8}
A cost graph crosses the vertical axis at \$15. Interpret it.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 9}
A population graph bends toward a ceiling. What does the ceiling suggest?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 10}
Data $3,6,12,24$. Linear or exponential trend?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 11}
A graph is horizontal at $y=50$. What is its rate of change?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 12}
Name two things a good graph description should include.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 13 --- Challenge}
App users climb steeply, then bend toward $1\,000\,000$. Predict cautiously and explain why linear extrapolation of the steep part fails.\workspace{3.2cm}
\end{qbox}

\clearpage
\sech{Question Recap \& Answer Key}

\begin{recapbox}
\begin{multicols}{2}
\small
\begin{enumerate}[leftmargin=*,itemsep=3pt]
  \item A graph rises then stays flat. Describe the trend.
  \item A distance–time line goes $(0,0)$ to $(4,200)$. Find the speed with units.
  \item Sales are $10,20,40,80$ over four years. Predict year $5$ and name the pattern.
  \item A graph falls and flattens. Is the size of the rate of change increasing or decreasing?
  \item A profit graph peaks at $(30,450)$. What does the peak mean?
  \item Sales rise by \$5000 each year from \$25\,000. Predict after $3$ more years.
  \item Why should a prediction far beyond the data be treated with caution?
  \item A cost graph crosses the vertical axis at \$15. Interpret it.
  \item A population graph bends toward a ceiling. What does the ceiling suggest?
  \item Data $3,6,12,24$. Linear or exponential trend?
  \item A graph is horizontal at $y=50$. What is its rate of change?
  \item Name two things a good graph description should include.
  \item App users climb steeply, then bend toward $1\,000\,000$. Predict cautiously and explain why linear extrapolation of the steep part fails.
\end{enumerate}
\end{multicols}
\end{recapbox}

\begin{keybox}
\begin{multicols}{2}
\small
\begin{enumerate}[leftmargin=*,itemsep=3pt]
  \item increases, then constant
  \item $200/4=50$ km/h
  \item $160$, exponential (doubling)
  \item decreasing
  \item maximum value $450$ at $30$
  \item $25\,000+3(5000)=\$40\,000$
  \item the trend may not continue (limits, new factors)
  \item a fixed starting cost of \$15
  \item a saturation limit (maximum size)
  \item exponential (constant ratio $2$)
  \item zero
  \item direction/trend, quantities with units, special points
  \item growth slows near the ceiling, so linear extrapolation over-predicts; approaches the $1\,000\,000$ limit
\end{enumerate}
\end{multicols}
\end{keybox}

\end{document}
