\documentclass[11pt]{article}
\input{_preamble}
\fancyhead[R]{\footnotesize MAP4C \textperiodcentered{} 2.5}
\begin{document}

\begin{center}
{\LARGE\bfseries\color{titleblue}Worksheet 2.5 --- Renting vs Owning Accommodation}\\[2pt]
{\small Grade 12 Foundations for College Mathematics (MAP4C) \textperiodcentered{} Unit 2: Personal Finance}
\end{center}

\begin{ideabox}
Compare the fixed and variable costs of renting and owning, test affordability against about $30\%$ of gross income, and weigh the financial and non-financial trade-offs.
\begin{itemize}[leftmargin=*,itemsep=2pt]
  \item Fixed vs variable housing costs.
  \item Affordability $\le$ about $30\%$ of gross income.
  \item Owning builds equity; renting adds flexibility.
\end{itemize}
\end{ideabox}

\begin{learnbox}
\lh{Fixed vs variable}Fixed costs (rent/mortgage, insurance, tax) stay steady; variable costs (utilities, repairs) change with use.
\lh{Affordability}A common guideline: total housing cost $\le$ about $30\%$ of gross monthly income.
\lh{Trade-offs}Owning builds equity but adds tax and repairs; renting costs less up front but builds none.
\end{learnbox}

\sech{Worked Examples}

\begin{example}{ex0}{Example 1 --- Cost of owning}
Mortgage $\$1288.60$, tax $\$250$, insurance $\$100$, maintenance $\$150$. Monthly cost?\soln $1288.60+250+100+150=\$1788.60$.
\end{example}

\begin{example}{ex1}{Example 2 --- Cost of renting}
Rent $\$1400$ plus tenant insurance $\$30$. Monthly cost?\soln $1400+30=\$1430$.
\end{example}

\begin{example}{ex2}{Example 3 --- Affordability}
Gross income $\$60\,000$/yr. Is $\$1789$ within the $30\%$ guideline?\soln $0.30(5000)=\$1500$; $\$1789>\$1500$ — over the guideline.
\end{example}

\begin{example}{ex3}{Example 4 --- Cost over a year}
How much more does owning ($\$1788.60$) cost than renting ($\$1430$) per year?\soln $358.60\times12\approx\$4303$.
\end{example}

\begin{example}{ex4}{Example 5 --- Trade-offs}
Give one financial and one non-financial reason to own.\soln Financial: builds equity. Non-financial: freedom to renovate (but you handle repairs).
\end{example}

\begin{example}{ex5}{Example 6 --- Variable cost}
Which is variable: property tax or utilities?\soln Utilities (they change with use).
\end{example}

\begin{example}{ex6}{Example 7 --- Affordability limit}
Gross income $\$48\,000$/yr. What is the $30\%$ monthly limit?\soln $0.30(4000)=\$1200$.
\end{example}

\begin{example}{ex7}{Example 8 --- Within the limit?}
Rent $\$1350$ + insurance $\$25$ — within the $\$1200$ limit?\soln $\$1375>\$1200$ — over.
\end{example}

\begin{example}{ex8}{Example 9 --- Equity adjustment}
Owning is $\$1789$/month, of which $\$289$ builds equity. What is the non-equity cost?\soln $1789-289=\$1500$/month.
\end{example}

\clearpage
\sech{Your Turn --- Show Your Work}
For each question, show all steps clearly in the space provided.

\begin{qbox}{Question 1}
Owning: mortgage $\$1100$, tax $\$200$, insurance $\$90$, maintenance $\$120$. Monthly cost?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 2}
Which is a variable cost: rent, property tax, or utilities?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 3}
Gross income $\$54\,000$/yr. What monthly cost is the $30\%$ limit?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 4}
Rent $\$1300$ + insurance $\$30$ — within a $\$1350$ limit?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 5}
Owning $\$1650$/month vs renting $\$1400$/month — yearly difference?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 6}
Name one financial advantage of owning.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 7}
Name one advantage of renting.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 8}
Gross income $\$72\,000$/yr — the $30\%$ monthly housing limit?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 9}
Owning: mortgage $\$1000$, tax $\$180$, insurance $\$85$, maintenance $\$110$. Monthly cost?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 10}
Is a down payment a fixed monthly cost?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 11}
Rent $\$1250$ + insurance $\$25$ per month — yearly total?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 12}
Why can owning be "cheaper" than the sticker cost suggests?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 13 --- Challenge}
Owning costs $\$1650$/month ($\$260$ builds equity); renting is $\$1400$. Compare the "true" extra cost after removing equity.\workspace{3cm}
\end{qbox}

\clearpage
\sech{Question Recap \& Answer Key}

\begin{recapbox}
\begin{multicols}{2}
\small
\begin{enumerate}[leftmargin=*,itemsep=3pt]
  \item Owning: mortgage $\$1100$, tax $\$200$, insurance $\$90$, maintenance $\$120$. Monthly cost?
  \item Which is a variable cost: rent, property tax, or utilities?
  \item Gross income $\$54\,000$/yr. What monthly cost is the $30\%$ limit?
  \item Rent $\$1300$ + insurance $\$30$ — within a $\$1350$ limit?
  \item Owning $\$1650$/month vs renting $\$1400$/month — yearly difference?
  \item Name one financial advantage of owning.
  \item Name one advantage of renting.
  \item Gross income $\$72\,000$/yr — the $30\%$ monthly housing limit?
  \item Owning: mortgage $\$1000$, tax $\$180$, insurance $\$85$, maintenance $\$110$. Monthly cost?
  \item Is a down payment a fixed monthly cost?
  \item Rent $\$1250$ + insurance $\$25$ per month — yearly total?
  \item Why can owning be "cheaper" than the sticker cost suggests?
  \item Owning costs $\$1650$/month ($\$260$ builds equity); renting is $\$1400$. Compare the "true" extra cost after removing equity.
\end{enumerate}
\end{multicols}
\end{recapbox}

\begin{keybox}
\begin{multicols}{2}
\small
\begin{enumerate}[leftmargin=*,itemsep=3pt]
  \item $1100+200+90+120=\$1510$
  \item utilities
  \item $0.30(4500)=\$1350$
  \item $\$1330<\$1350$ — yes
  \item $(1650-1400)\times12=\$3000$
  \item builds equity
  \item flexibility to move (no repairs, lower up-front cost)
  \item $0.30(6000)=\$1800$
  \item $1000+180+85+110=\$1375$
  \item no — it is a one-time up-front cost
  \item $1275\times12=\$15\,300$
  \item part of the mortgage payment builds equity (savings, not spending)
  \item non-equity owning $1650-260=\$1390$; vs rent $\$1400$, about $\$10$/month less
\end{enumerate}
\end{multicols}
\end{keybox}

\end{document}
