\documentclass[11pt]{article}
\input{_preamble}
\fancyhead[R]{\footnotesize MCT4C \textperiodcentered{} 2.2}
\begin{document}

\begin{center}
{\LARGE\bfseries\color{titleblue}Worksheet 2.2 --- Graphs of Polynomial Functions}\\[2pt]
{\small Grade 12 Mathematics for College Technology (MCT4C) \textperiodcentered{} Unit 2: Polynomial Functions}
\end{center}

\begin{ideabox}
Degree parity and the leading coefficient's sign set the \emph{end behaviour}; a degree-$n$ polynomial has at most $n$ $x$-intercepts and $n-1$ turning points.
\begin{itemize}[leftmargin=*,itemsep=2pt]
  \item Even degree: ends match; odd degree: ends oppose.
  \item $+$ leading rises right; $-$ leading falls right.
  \item At most $n$ intercepts, $n-1$ turns.
\end{itemize}
\end{ideabox}

\begin{learnbox}
\lh{End behaviour}The degree's parity and the leading coefficient's sign fix the ends: even degrees send both ends the same way, odd degrees send them opposite ways.
\lh{Intercepts and turns}A degree-$n$ polynomial has at most $n$ $x$-intercepts and $n-1$ turning points, so the degree caps the graph's complexity.
\lh{Zeros and multiplicity}At a zero of odd multiplicity the graph crosses the axis; at even multiplicity it touches and turns back.
\end{learnbox}

\sech{Worked Examples}

\begin{example}{ex0}{Example 1 --- Odd, positive}
End behaviour of $y=2x^3-5x+1$?\soln Down-left, up-right.
\end{example}

\begin{example}{ex1}{Example 2 --- Even, negative}
End behaviour of $y=-x^4+3x^2$?\soln Down on both ends.
\end{example}

\begin{example}{ex2}{Example 3 --- Count features}
Max $x$-intercepts and turning points of a quartic?\soln $4$ and $3$.
\end{example}

\begin{example}{ex3}{Example 4 --- Reason from graph}
A graph rises on both ends. Degree and leading sign?\soln Even degree, positive leading coefficient.
\end{example}

\begin{example}{ex4}{Example 5 --- Odd, negative}
End behaviour of $y=-x^5+2x^2$?\soln Up-left, down-right.
\end{example}

\begin{example}{ex5}{Example 6 --- Cubic ends}
End behaviour of $y=x^3$?\soln Down-left, up-right.
\end{example}

\begin{example}{ex6}{Example 7 --- Quartic ends}
End behaviour of $y=x^4$?\soln Up on both ends.
\end{example}

\begin{example}{ex7}{Example 8 --- Turning points}
Max turning points of a degree-6 polynomial?\soln $5$.
\end{example}

\begin{example}{ex8}{Example 9 --- Falls both ends}
A graph falls on both ends. Degree and sign?\soln Even degree, negative leading coefficient.
\end{example}

\clearpage
\sech{Your Turn --- Show Your Work}
For each question, show all steps clearly in the space provided.

\begin{qbox}{Question 1}
End behaviour of $y=x^4-2x^2$?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 2}
End behaviour of $y=-2x^3+x$?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 3}
Max $x$-intercepts of a cubic?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 4}
A graph falls both ends — degree and leading sign?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 5}
Max turning points of a degree-6 polynomial?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 6}
End behaviour of $y=3x^5$?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 7}
End behaviour of $y=-x^6$?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 8}
Max $x$-intercepts of a quartic?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 9}
A graph rises both ends — degree parity?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 10}
End behaviour of $y=-5x^3+2x^2-1$?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 11}
Max turning points of a cubic?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 12}
End behaviour of $y=2x^4-x$?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 13 --- Challenge}
A degree-5 graph with negative leading coefficient — describe its ends.\workspace{2.6cm}
\end{qbox}

\clearpage
\sech{Question Recap \& Answer Key}

\begin{recapbox}
\begin{multicols}{2}
\small
\begin{enumerate}[leftmargin=*,itemsep=3pt]
  \item End behaviour of $y=x^4-2x^2$?
  \item End behaviour of $y=-2x^3+x$?
  \item Max $x$-intercepts of a cubic?
  \item A graph falls both ends — degree and leading sign?
  \item Max turning points of a degree-6 polynomial?
  \item End behaviour of $y=3x^5$?
  \item End behaviour of $y=-x^6$?
  \item Max $x$-intercepts of a quartic?
  \item A graph rises both ends — degree parity?
  \item End behaviour of $y=-5x^3+2x^2-1$?
  \item Max turning points of a cubic?
  \item End behaviour of $y=2x^4-x$?
  \item A degree-5 graph with negative leading coefficient — describe its ends.
\end{enumerate}
\end{multicols}
\end{recapbox}

\begin{keybox}
\begin{multicols}{2}
\small
\begin{enumerate}[leftmargin=*,itemsep=3pt]
  \item up both ends
  \item up-left, down-right
  \item $3$
  \item even degree, negative LC
  \item $5$
  \item down-left, up-right
  \item down both ends
  \item $4$
  \item even
  \item up-left, down-right
  \item $2$
  \item up both ends
  \item up-left, down-right
\end{enumerate}
\end{multicols}
\end{keybox}

\end{document}
