\documentclass[11pt]{article}
\input{_preamble}
\fancyhead[R]{\footnotesize MCT4C \textperiodcentered{} 2.8}
\begin{document}

\begin{center}
{\LARGE\bfseries\color{titleblue}Worksheet 2.8 --- Applications of Polynomial Functions}\\[2pt]
{\small Grade 12 Mathematics for College Technology (MCT4C) \textperiodcentered{} Unit 2: Polynomial Functions}
\end{center}

\begin{ideabox}
Model a situation with a polynomial (usually a volume or height), set the equation, solve by factoring, and keep only physically sensible roots (positive lengths).
\begin{itemize}[leftmargin=*,itemsep=2pt]
  \item Translate the context into a polynomial.
  \item Solve by factoring.
  \item Discard out-of-range roots.
\end{itemize}
\end{ideabox}

\begin{learnbox}
\lh{Modelling with polynomials}Volumes, projectile heights, and design quantities often form a polynomial; define the variable and write the relation.
\lh{Solving in context}Factor and solve, then keep only the physically sensible roots — lengths and times must be positive.
\lh{Interpreting the answer}Match the question to the feature: a maximum points to a turning point, a landing or break-even to a zero.
\end{learnbox}

\sech{Worked Examples}

\begin{example}{ex0}{Example 1 --- Open box}
Squares of side $x$ are cut from a $12\times12$ sheet. For what $x$ is the volume $128$?\soln $V=x(12-2x)^2$; $x=2$ gives $2(8)^2=128$ ($0<x<6$).
\end{example}

\begin{example}{ex1}{Example 2 --- Projectile}
$h(t)=-5t^2+30t$. When does it land?\soln $t(-5t+30)=0\Rightarrow t=6$ s.
\end{example}

\begin{example}{ex2}{Example 3 --- Volume from factors}
$V=x^3+4x^2+3x$; factor and find $V$ at $x=2$.\soln $x(x+1)(x+3)$; at $2$: $2\cdot3\cdot5=30$.
\end{example}

\begin{example}{ex3}{Example 4 --- Break-even}
Solve $x^3-7x-6=0$ for the positive root.\soln $(x-3)(x+1)(x+2)=0\Rightarrow x=3$.
\end{example}

\begin{example}{ex4}{Example 5 --- Beam deflection}
$d(x)=x^3-12x$. Where is $d=0$?\soln $x(x^2-12)=0\Rightarrow x=0,\ 2\sqrt3$.
\end{example}

\begin{example}{ex5}{Example 6 --- Landing time}
$h(t)=-5t^2+20t$. When does it land?\soln $t(-5t+20)=0\Rightarrow t=4$ s.
\end{example}

\begin{example}{ex6}{Example 7 --- Box dimensions}
A box has volume $x(x-1)(x+2)$; find $V$ at $x=3$.\soln $3\cdot2\cdot5=30$.
\end{example}

\begin{example}{ex7}{Example 8 --- Design equation}
Solve $x^3-9x=0$ for the positive length.\soln $x=3$.
\end{example}

\begin{example}{ex8}{Example 9 --- Profit roots}
Solve $x^3-x^2-4x+4=0$; give the positive roots.\soln $x=1,2$.
\end{example}

\clearpage
\sech{Your Turn --- Show Your Work}
For each question, show all steps clearly in the space provided.

\begin{qbox}{Question 1}
$h(t)=-5t^2+20t$. When does it land?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 2}
A box $x(x-1)(x+2)$; find $V$ at $x=3$.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 3}
Solve $x^3-9x=0$ (positive length).\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 4}
$V=x(10-2x)^2$; is $V=72$ at $x=1$?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 5}
Solve $x^3-2x^2-5x+6=0$; positive roots?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 6}
$h(t)=-5t^2+40t$. When does it land?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 7}
$V=x^3+5x^2+6x$; find $V$ at $x=1$.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 8}
Solve $x^3-16x=0$ (positive length).\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 9}
A box $x(8-2x)^2=72$; check $x=1$.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 10}
Solve $x^3-4x=0$ (positive length).\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 11}
$d(x)=x^3-9x$; where is $d=0$ (positive)?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 12}
$V=x(x+2)(x+4)$; find $V$ at $x=2$.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 13 --- Challenge}
A $16\times16$ sheet, cut $x$: is $V=x(16-2x)^2=768$ at $x=2$?\workspace{3.2cm}
\end{qbox}

\clearpage
\sech{Question Recap \& Answer Key}

\begin{recapbox}
\begin{multicols}{2}
\small
\begin{enumerate}[leftmargin=*,itemsep=3pt]
  \item $h(t)=-5t^2+20t$. When does it land?
  \item A box $x(x-1)(x+2)$; find $V$ at $x=3$.
  \item Solve $x^3-9x=0$ (positive length).
  \item $V=x(10-2x)^2$; is $V=72$ at $x=1$?
  \item Solve $x^3-2x^2-5x+6=0$; positive roots?
  \item $h(t)=-5t^2+40t$. When does it land?
  \item $V=x^3+5x^2+6x$; find $V$ at $x=1$.
  \item Solve $x^3-16x=0$ (positive length).
  \item A box $x(8-2x)^2=72$; check $x=1$.
  \item Solve $x^3-4x=0$ (positive length).
  \item $d(x)=x^3-9x$; where is $d=0$ (positive)?
  \item $V=x(x+2)(x+4)$; find $V$ at $x=2$.
  \item A $16\times16$ sheet, cut $x$: is $V=x(16-2x)^2=768$ at $x=2$?
\end{enumerate}
\end{multicols}
\end{recapbox}

\begin{keybox}
\begin{multicols}{2}
\small
\begin{enumerate}[leftmargin=*,itemsep=3pt]
  \item $t=4$ s
  \item $30$
  \item $x=3$
  \item $1(8)^2=64\ne72$: no
  \item $x=1,3$
  \item $t=8$ s
  \item $1\cdot3\cdot4=12$
  \item $x=4$
  \item $1(6)^2=36\ne72$: no
  \item $x=2$
  \item $x=3$
  \item $2\cdot4\cdot6=48$
  \item $2(12)^2=288\ne768$: no
\end{enumerate}
\end{multicols}
\end{keybox}

\end{document}
