\documentclass[11pt]{article}
\input{_preamble}
\fancyhead[R]{\footnotesize MPM2D \textperiodcentered{} 6.6}
\begin{document}

\begin{center}
{\LARGE\bfseries\color{titleblue}Worksheet 6.6 --- Solving Right Triangles \& Applications}\\[2pt]
{\small Grade 10 Academic (MPM2D) \textperiodcentered{} Unit 6: Trigonometry}
\end{center}

\begin{ideabox}
To \textbf{solve} a right triangle is to find all unknown sides and angles. Real problems use angles of elevation and depression.
\begin{itemize}[leftmargin=*,itemsep=2pt]
  \item With one angle and one side, find the rest using SOH CAH TOA and the Pythagorean theorem.
  \item The two acute angles sum to $90^\circ$.
  \item \textbf{Angle of elevation} is measured up from the horizontal; \textbf{depression} is measured down.
  \item Draw and label a diagram before choosing a ratio.
\end{itemize}
\end{ideabox}

\begin{learnbox}
\lh{Solve the whole triangle}\emph{Solving} a right triangle means finding every unknown side and angle, using the trig ratios and the fact that the angles sum to $180^\circ$.
\lh{Elevation and depression}Real problems use the \textbf{angle of elevation} (looking up) or \textbf{depression} (looking down) from the horizontal to build the triangle.
\lh{Draw, then decide}Sketch and label the triangle first, then choose the ratio or theorem each step requires.
\end{learnbox}

\sech{Worked Examples}

\begin{example}{ex0}{Example 1 --- Solve from angle + hyp}
Solve the right triangle with $\theta=30^\circ$ and hyp $=10$.
\soln Opp $=10\sin30^\circ=5$; adj $=10\cos30^\circ\approx8.7$; other angle $=60^\circ.$
\end{example}

\begin{example}{ex1}{Example 2 --- Solve from two legs}
Solve the right triangle with legs $3$ and $4$.
\soln Hyp $=5$; angles $\tan^{-1}(\tfrac34)\approx36.9^\circ$ and $53.1^\circ.$
\end{example}

\begin{example}{ex2}{Example 3 --- Angle of elevation}
From $50$ m away, the angle of elevation to a tower's top is $40^\circ$. Find its height.
\soln $h=50\tan40^\circ\approx42.0$ m.
\end{example}

\begin{example}{ex3}{Example 4 --- Angle of depression}
From an $80$ m cliff, the angle of depression to a boat is $25^\circ$. Find the boat's distance from the base.
\soln $d=\dfrac{80}{\tan25^\circ}\approx171.6$ m.
\end{example}

\begin{example}{ex4}{Example 5 --- Ladder}
A $5$ m ladder rests with its base $1.5$ m from a wall. Find the angle with the ground and the height reached.
\soln Angle $=\cos^{-1}\!\left(\tfrac{1.5}{5}\right)\approx72.5^\circ$; height $=\sqrt{5^2-1.5^2}\approx4.8$ m.
\end{example}

\begin{example}{ex5}{Example 6 --- Ramp angle}
A ramp rises $2$ m over a $10$ m horizontal run. Find its angle.
\soln $\theta=\tan^{-1}\!\left(\tfrac{2}{10}\right)\approx11.3^\circ.$
\end{example}

\begin{example}{ex6}{Example 7 --- Kite}
A kite string $60$ m long makes a $50^\circ$ angle with the ground. Find the kite's height.
\soln $h=60\sin50^\circ\approx46.0$ m.
\end{example}

\begin{example}{ex7}{Example 8 --- Elevation to a building}
From $100$ m away, the angle of elevation to a building's top is $35^\circ$. Find its height.
\soln $h=100\tan35^\circ\approx70.0$ m.
\end{example}

\begin{example}{ex8}{Example 9 --- Two-step (harder)}
A $12$ m flagpole casts a shadow when the sun's elevation is $58^\circ$. Find the shadow length, then the angle when the shadow is $20$ m.
\soln Shadow $=\dfrac{12}{\tan58^\circ}\approx7.5$ m. When the shadow is $20$ m: $\theta=\tan^{-1}\!\left(\tfrac{12}{20}\right)\approx31.0^\circ.$
\end{example}

\clearpage
\sech{Your Turn --- Show Your Work}
For each question, show all steps clearly in the space provided.

\begin{qbox}{Question 1}
Solve the right triangle with $\theta=40^\circ$ and hyp $=20$.\workspace{2.6cm}
\end{qbox}

\begin{qbox}{Question 2}
Solve the right triangle with legs $6$ and $8$.\workspace{2.6cm}
\end{qbox}

\begin{qbox}{Question 3}
From $30$ m away, the angle of elevation to a tree top is $50^\circ$. Find its height.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 4}
From a $60$ m tower, the angle of depression to a car is $35^\circ$. Find its distance from the base.\workspace{2.6cm}
\end{qbox}

\begin{qbox}{Question 5}
A $4$ m ladder's base is $1$ m from a wall. Find the angle with the ground.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 6}
A ramp rises $3$ m over a $12$ m run. Find its angle.\workspace{2.2cm}
\end{qbox}

\begin{qbox}{Question 7}
A kite string $80$ m long makes a $40^\circ$ angle. Find the kite's height.\workspace{2.2cm}
\end{qbox}

\begin{qbox}{Question 8}
From $120$ m away, the elevation to a building top is $28^\circ$. Find its height.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 9}
Solve the right triangle with $\theta=55^\circ$ and adjacent $14$.\workspace{2.6cm}
\end{qbox}

\begin{qbox}{Question 10}
A $10$ m wire is anchored $4$ m from the base of a pole. Find the angle the wire makes with the ground.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 11}
From a $45$ m lighthouse, the depression to a boat is $20^\circ$. Find the boat's distance.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 12}
A slide is $5$ m long and meets the ground at $30^\circ$. Find its height.\workspace{2.2cm}
\end{qbox}

\begin{qbox}{Question 13 --- Challenge}
A $15$ m pole casts a shadow when the sun's elevation is $62^\circ$. Find the shadow, then the elevation when the shadow is $25$ m.\workspace{3.2cm}
\end{qbox}

\clearpage
\sech{Question Recap \& Answer Key}

\begin{recapbox}
\begin{multicols}{2}
\small
\begin{enumerate}[leftmargin=*,itemsep=3pt]
  \item Solve the right triangle with $\theta=40^\circ$ and hyp $=20$.
  \item Solve the right triangle with legs $6$ and $8$.
  \item From $30$ m away, the angle of elevation to a tree top is $50^\circ$. Find its height.
  \item From a $60$ m tower, the angle of depression to a car is $35^\circ$. Find its distance from the base.
  \item A $4$ m ladder's base is $1$ m from a wall. Find the angle with the ground.
  \item A ramp rises $3$ m over a $12$ m run. Find its angle.
  \item A kite string $80$ m long makes a $40^\circ$ angle. Find the kite's height.
  \item From $120$ m away, the elevation to a building top is $28^\circ$. Find its height.
  \item Solve the right triangle with $\theta=55^\circ$ and adjacent $14$.
  \item A $10$ m wire is anchored $4$ m from the base of a pole. Find the angle the wire makes with the ground.
  \item From a $45$ m lighthouse, the depression to a boat is $20^\circ$. Find the boat's distance.
  \item A slide is $5$ m long and meets the ground at $30^\circ$. Find its height.
  \item A $15$ m pole casts a shadow when the sun's elevation is $62^\circ$. Find the shadow, then the elevation when the shadow is $25$ m.
\end{enumerate}
\end{multicols}
\end{recapbox}

\begin{keybox}
\begin{multicols}{2}
\small
\begin{enumerate}[leftmargin=*,itemsep=3pt]
  \item Opp $=20\sin40^\circ\approx12.9$; adj $\approx15.3$; other angle $50^\circ$.
  \item Hyp $10$; angles $\approx36.9^\circ$ and $53.1^\circ$.
  \item $30\tan50^\circ\approx35.8$ m.
  \item $\tfrac{60}{\tan35^\circ}\approx85.7$ m.
  \item $\cos^{-1}(0.25)\approx75.5^\circ$.
  \item $\tan^{-1}(0.25)\approx14.0^\circ$.
  \item $80\sin40^\circ\approx51.4$ m.
  \item $120\tan28^\circ\approx63.8$ m.
  \item Opp $=14\tan55^\circ\approx20.0$; hyp $\approx24.4$; other angle $35^\circ$.
  \item $\cos^{-1}(0.4)\approx66.4^\circ$.
  \item $\tfrac{45}{\tan20^\circ}\approx123.6$ m.
  \item $5\sin30^\circ=2.5$ m.
  \item Shadow $=\tfrac{15}{\tan62^\circ}\approx8.0$ m; then $\tan^{-1}(\tfrac{15}{25})\approx31.0^\circ$.
\end{enumerate}
\end{multicols}
\end{keybox}

\end{document}
