\documentclass[11pt]{article}
\input{_preamble}
\fancyhead[R]{\footnotesize MCV4U \textperiodcentered{} 5.3}
\begin{document}

\begin{center}
{\LARGE\bfseries\color{titleblue}Worksheet 5.3 --- Kinematics: Velocity \& Acceleration}\\[2pt]
{\small Grade 12 Calculus and Vectors (MCV4U) \textperiodcentered{} Unit 5: Applications of Derivatives}
\end{center}

\begin{ideabox}
$v=s'$, $a=v'=s''$, speed $=|v|$. At rest when $v=0$; speeding up when $v,a$ share a sign.
\begin{itemize}[leftmargin=*,itemsep=2pt]
  \item $v(t)=s'(t)$, $a(t)=s''(t)$.
  \item At rest: $v=0$; direction: sign of $v$.
  \item Speeding up: $v,a$ same sign; slowing: opposite.
\end{itemize}
\end{ideabox}

\begin{learnbox}
\lh{Position, velocity, acceleration}If $s(t)$ is position, then velocity is $v(t)=s'(t)$ and acceleration is $a(t)=v'(t)=s''(t)$ — each a derivative of the last.
\lh{Reading motion}Velocity's sign gives direction; where $v=0$ the object is momentarily at rest and may turn around.
\lh{Speeding up or slowing}The object speeds up when velocity and acceleration share a sign, and slows when their signs differ.
\end{learnbox}

\sech{Worked Examples}

\begin{example}{ex0}{Example 1 --- v and a}
For $s(t)=t^2-4t$, find $v$ and $a$.\soln $v=2t-4$, $a=2$. Velocity is the slope of position; at rest at the lowest point ($t=2$):\eplot{-1}{5}{-5}{6}{\addplot[exblue,very thick,domain=0:4.5]{x^2-4*x};\addplot[qorange,very thick,domain=0:4.5]{2*x-4};\addplot[red,only marks,mark=*,mark size=1.6pt] coordinates {(2,-4)};}
\end{example}

\begin{example}{ex1}{Example 2 --- At rest}
When is $s(t)=t^2-4t$ at rest?\soln $v=2t-4=0\Rightarrow t=2$ s.
\end{example}

\begin{example}{ex2}{Example 3 --- Multiple rest times}
When is $s(t)=t^3-6t^2+9t$ at rest?\soln $v=3t^2-12t+9=3(t-1)(t-3)=0\Rightarrow t=1,3$.
\end{example}

\begin{example}{ex3}{Example 4 --- Direction \& accel}
Describe the motion of $s(t)=t^3-6t^2+9t$ at $t=0$.\soln $v(0)=9>0$ (moving right); $a=6t-12$, $a(0)=-12<0$ (decelerating).
\end{example}

\begin{example}{ex4}{Example 5 --- Speeding up?}
Is $s(t)=t^3-6t^2+9t$ speeding up at $t=4$?\soln $v(4)=9>0$, $a(4)=12>0$; same sign ⇒ speeding up.
\end{example}

\begin{example}{ex5}{Example 6 --- Velocity}
For $s(t)=t^2-6t$, find $v(t)$.\soln $2t-6$.
\end{example}

\begin{example}{ex6}{Example 7 --- At rest}
When is $s(t)=t^2-6t$ at rest?\soln $2t-6=0\Rightarrow t=3$.
\end{example}

\begin{example}{ex7}{Example 8 --- Acceleration}
For $s(t)=t^3-3t^2$, find $a(t)$.\soln $v=3t^2-6t$, $a=6t-6$.
\end{example}

\begin{example}{ex8}{Example 9 --- Max height}
A projectile has $h(t)=-5t^2+20t$. When is it highest?\soln $v=-10t+20=0\Rightarrow t=2$ s.
\end{example}

\clearpage
\sech{Your Turn --- Show Your Work}
For each question, show all steps clearly in the space provided.

\begin{qbox}{Question 1}
For $s(t)=t^2-6t$, find $v(t)$.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 2}
When is $s(t)=t^2-6t$ at rest?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 3}
For $s(t)=t^3-3t^2$, find $a(t)$.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 4}
A projectile has $h(t)=-5t^2+30t$. When is it highest?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 5}
If $v>0$ and $a<0$, is the object speeding up or slowing down?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 6}
For $s(t)=2t^2-8t$, find $v(t)$ and $a(t)$.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 7}
When is $s(t)=t^3-6t^2+9t$ at rest?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 8}
For $s(t)=t^3-3t^2$, when is the object at rest?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 9}
At $t=0$, $s(t)=t^3-6t^2+9t$: which direction is it moving?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 10}
Speed of $s(t)=t^2-4t$ at $t=0$?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 11}
If $v<0$ and $a<0$, is the object speeding up or slowing down?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 12}
Acceleration of $s(t)=t^2-4t$ (constant)?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 13 --- Challenge}
For $s(t)=t^3-6t^2+9t$, determine whether it is speeding up or slowing down at $t=2$.\workspace{3cm}
\end{qbox}

\clearpage
\sech{Question Recap \& Answer Key}

\begin{recapbox}
\begin{multicols}{2}
\small
\begin{enumerate}[leftmargin=*,itemsep=3pt]
  \item For $s(t)=t^2-6t$, find $v(t)$.
  \item When is $s(t)=t^2-6t$ at rest?
  \item For $s(t)=t^3-3t^2$, find $a(t)$.
  \item A projectile has $h(t)=-5t^2+30t$. When is it highest?
  \item If $v>0$ and $a<0$, is the object speeding up or slowing down?
  \item For $s(t)=2t^2-8t$, find $v(t)$ and $a(t)$.
  \item When is $s(t)=t^3-6t^2+9t$ at rest?
  \item For $s(t)=t^3-3t^2$, when is the object at rest?
  \item At $t=0$, $s(t)=t^3-6t^2+9t$: which direction is it moving?
  \item Speed of $s(t)=t^2-4t$ at $t=0$?
  \item If $v<0$ and $a<0$, is the object speeding up or slowing down?
  \item Acceleration of $s(t)=t^2-4t$ (constant)?
  \item For $s(t)=t^3-6t^2+9t$, determine whether it is speeding up or slowing down at $t=2$.
\end{enumerate}
\end{multicols}
\end{recapbox}

\begin{keybox}
\begin{multicols}{2}
\small
\begin{enumerate}[leftmargin=*,itemsep=3pt]
  \item $2t-6$
  \item $t=3$
  \item $a=6t-6$
  \item $v=-10t+30=0\Rightarrow t=3$ s
  \item slowing down
  \item $v=4t-8$, $a=4$
  \item $t=1,3$
  \item $v=3t^2-6t=0\Rightarrow t=0,2$
  \item right ($v(0)=9>0$)
  \item $|v(0)|=|-4|=4$
  \item speeding up
  \item $2$
  \item $v(2)=-3<0$, $a(2)=0$ — at an instant of zero acceleration (neither)
\end{enumerate}
\end{multicols}
\end{keybox}

\end{document}
