\documentclass[11pt]{article}
\input{_preamble}
\fancyhead[R]{\footnotesize MCV4U \textperiodcentered{} 6.1}
\begin{document}

\begin{center}
{\LARGE\bfseries\color{titleblue}Worksheet 6.1 --- Introduction to Vectors}\\[2pt]
{\small Grade 12 Calculus and Vectors (MCV4U) \textperiodcentered{} Unit 6: Geometry and Algebra of Vectors}
\end{center}

\begin{ideabox}
A vector has magnitude and direction (velocity, force); a scalar has magnitude only (speed, mass). Equal vectors share magnitude and direction; $-\vec v$ reverses direction.
\begin{itemize}[leftmargin=*,itemsep=2pt]
  \item Vector = magnitude + direction; scalar = magnitude only.
  \item Equal vectors: same magnitude and direction (position-free).
  \item Directions via bearings ($N30^\circ E$, $030^\circ$).
\end{itemize}
\end{ideabox}

\begin{learnbox}
\lh{Magnitude and direction}A \textbf{vector} has both size and direction, unlike a scalar. Displacement, velocity, and force are vector quantities.
\lh{Representing vectors}Draw a vector as a directed arrow; two vectors are equal when they share magnitude and direction, wherever they are placed.
\lh{Special vectors}The \textbf{zero vector} has no length, and the opposite of a vector reverses its direction.
\end{learnbox}

\sech{Worked Examples}

\begin{example}{ex0}{Example 1 --- Vector or scalar}
Classify velocity, speed, force, temperature, displacement.\soln Vectors: velocity, force, displacement. Scalars: speed, temperature.
\end{example}

\begin{example}{ex1}{Example 2 --- Magnitude \& direction}
A plane flies 500 km at $N60^\circ E$.\soln Magnitude $500$ km; direction $N60^\circ E$.
\end{example}

\begin{example}{ex2}{Example 3 --- Equal vectors}
When are two displacement vectors equal?\soln When they share magnitude and direction — location is irrelevant.
\end{example}

\begin{example}{ex3}{Example 4 --- Opposite vector}
If $\vec v$ is 8 m/s east, describe $-\vec v$.\soln 8 m/s west.
\end{example}

\begin{example}{ex4}{Example 5 --- Bearing}
Express "$40^\circ$ east of north" as a true bearing.\soln $040^\circ$.
\end{example}

\begin{example}{ex5}{Example 6 --- Vector or scalar}
Is acceleration a vector or scalar?\soln Vector.
\end{example}

\begin{example}{ex6}{Example 7 --- Vector or scalar}
Is mass a vector or scalar?\soln Scalar.
\end{example}

\begin{example}{ex7}{Example 8 --- Opposite}
Describe $-\vec v$ if $\vec v$ is 10 N north.\soln 10 N south.
\end{example}

\begin{example}{ex8}{Example 9 --- Bearing}
Write $N75^\circ E$ as a true bearing.\soln $075^\circ$.
\end{example}

\clearpage
\sech{Your Turn --- Show Your Work}
For each question, show all steps clearly in the space provided.

\begin{qbox}{Question 1}
Is force a vector or a scalar?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 2}
Is temperature a vector or a scalar?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 3}
Describe $-\vec v$ if $\vec v$ is 6 m/s west.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 4}
Two vectors have equal magnitude, opposite direction. Equal?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 5}
Write $N20^\circ E$ as a true bearing.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 6}
Is displacement a vector or a scalar?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 7}
Is speed a vector or a scalar?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 8}
Describe $-\vec v$ if $\vec v$ is 12 N east.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 9}
What two things define a vector?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 10}
Write $N50^\circ E$ as a true bearing.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 11}
Is energy a vector or a scalar?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 12}
Does the position of a vector affect whether two are equal?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 13 --- Challenge}
A ship sails 30 km at a bearing of $120^\circ$. State its magnitude and describe the bearing in N–E form.\workspace{2.6cm}
\end{qbox}

\clearpage
\sech{Question Recap \& Answer Key}

\begin{recapbox}
\begin{multicols}{2}
\small
\begin{enumerate}[leftmargin=*,itemsep=3pt]
  \item Is force a vector or a scalar?
  \item Is temperature a vector or a scalar?
  \item Describe $-\vec v$ if $\vec v$ is 6 m/s west.
  \item Two vectors have equal magnitude, opposite direction. Equal?
  \item Write $N20^\circ E$ as a true bearing.
  \item Is displacement a vector or a scalar?
  \item Is speed a vector or a scalar?
  \item Describe $-\vec v$ if $\vec v$ is 12 N east.
  \item What two things define a vector?
  \item Write $N50^\circ E$ as a true bearing.
  \item Is energy a vector or a scalar?
  \item Does the position of a vector affect whether two are equal?
  \item A ship sails 30 km at a bearing of $120^\circ$. State its magnitude and describe the bearing in N–E form.
\end{enumerate}
\end{multicols}
\end{recapbox}

\begin{keybox}
\begin{multicols}{2}
\small
\begin{enumerate}[leftmargin=*,itemsep=3pt]
  \item vector
  \item scalar
  \item 6 m/s east
  \item no (opposites)
  \item $020^\circ$
  \item vector
  \item scalar
  \item 12 N west
  \item magnitude and direction
  \item $050^\circ$
  \item scalar
  \item no
  \item 30 km; $120^\circ=S60^\circ E$
\end{enumerate}
\end{multicols}
\end{keybox}

\end{document}
