\documentclass[11pt]{article}
\input{_preamble}
\fancyhead[R]{\footnotesize MCV4U \textperiodcentered{} 6.2}
\begin{document}

\begin{center}
{\LARGE\bfseries\color{titleblue}Worksheet 6.2 --- Vector Operations (Geometric)}\\[2pt]
{\small Grade 12 Calculus and Vectors (MCV4U) \textperiodcentered{} Unit 6: Geometry and Algebra of Vectors}
\end{center}

\begin{ideabox}
Add tip-to-tail (triangle law); subtract via $\vec a-\vec b=\vec a+(-\vec b)$; scale by $k$ to change length (and flip if $k<0$).
\begin{itemize}[leftmargin=*,itemsep=2pt]
  \item Addition: tip-to-tail; resultant from first tail to last tip.
  \item Subtraction: $\vec a-\vec b=\vec a+(-\vec b)$.
  \item $k\vec v$ has length $|k||\vec v|$; reversed if $k<0$.
\end{itemize}
\end{ideabox}

\begin{learnbox}
\lh{Adding vectors}Add vectors \textbf{tip-to-tail}: the resultant runs from the first tail to the last tip. The parallelogram rule gives the same result.
\lh{Scalar multiples}Multiplying by a scalar scales the length and, if negative, reverses the direction while keeping the line of action.
\lh{Subtraction}Subtracting a vector means adding its opposite, which the triangle of vectors makes easy to picture.
\end{learnbox}

\sech{Worked Examples}

\begin{example}{ex0}{Example 1 --- Perpendicular resultant}
Add 3 east and 4 north; find the resultant.\soln $\sqrt{3^2+4^2}=5$, direction $\tan^{-1}\tfrac43\approx53^\circ$ N of E:\begin{center}\begin{tikzpicture}[scale=0.7]\draw[->,thick,exblue](0,0)--(3,0)node[midway,below]{$3$ E};\draw[->,thick,qorange](3,0)--(3,4)node[midway,right]{$4$ N};\draw[->,very thick,red](0,0)--(3,4)node[midway,above left]{$R=5$};\end{tikzpicture}\end{center}
\end{example}

\begin{example}{ex1}{Example 2 --- Triangle law}
How do you draw $\vec a+\vec b$?\soln Tail of $\vec b$ at the tip of $\vec a$; resultant from $\vec a$'s tail to $\vec b$'s tip.
\end{example}

\begin{example}{ex2}{Example 3 --- Scalar multiple}
Describe $3\vec v$ and $-2\vec v$.\soln $3\vec v$: three times as long, same direction; $-2\vec v$: twice as long, opposite.
\end{example}

\begin{example}{ex3}{Example 4 --- Subtraction}
How do you draw $\vec a-\vec b$?\soln Add $\vec a$ and $-\vec b$ tip-to-tail.
\end{example}

\begin{example}{ex4}{Example 5 --- Resultant force}
Forces 6 N east and 8 N north. Find the resultant.\soln $\sqrt{36+64}=10$ N at $\tan^{-1}\tfrac86\approx53^\circ$ N of E.
\end{example}

\begin{example}{ex5}{Example 6 --- Perpendicular resultant}
Resultant magnitude of 5 east and 12 north?\soln $\sqrt{25+144}=13$.
\end{example}

\begin{example}{ex6}{Example 7 --- Scalar multiple}
Describe $-\tfrac12\vec v$.\soln Half the length, opposite direction.
\end{example}

\begin{example}{ex7}{Example 8 --- Perpendicular resultant}
Resultant magnitude of 9 east and 12 north?\soln $\sqrt{81+144}=15$.
\end{example}

\begin{example}{ex8}{Example 9 --- Direction}
Direction of the resultant of 3 east, 4 north?\soln $\tan^{-1}\tfrac43\approx53^\circ$ north of east.
\end{example}

\clearpage
\sech{Your Turn --- Show Your Work}
For each question, show all steps clearly in the space provided.

\begin{qbox}{Question 1}
Resultant magnitude of 5 east and 12 north?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 2}
Describe $4\vec v$.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 3}
Describe $-\tfrac13\vec v$.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 4}
Resultant magnitude of 8 east and 15 north?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 5}
What law adds vectors tip-to-tail?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 6}
Resultant magnitude of 6 east and 8 north?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 7}
Describe $-3\vec v$.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 8}
Resultant magnitude of 7 east and 24 north?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 9}
How do you express $\vec a-\vec b$ as an addition?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 10}
Direction (N of E) of the resultant of 5 east, 5 north?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 11}
Resultant magnitude of 9 east and 40 north?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 12}
Length of $5\vec v$ if $|\vec v|=3$?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 13 --- Challenge}
Two forces, 10 N east and 24 N north, act on an object. Find the magnitude and direction of the resultant.\workspace{3cm}
\end{qbox}

\clearpage
\sech{Question Recap \& Answer Key}

\begin{recapbox}
\begin{multicols}{2}
\small
\begin{enumerate}[leftmargin=*,itemsep=3pt]
  \item Resultant magnitude of 5 east and 12 north?
  \item Describe $4\vec v$.
  \item Describe $-\tfrac13\vec v$.
  \item Resultant magnitude of 8 east and 15 north?
  \item What law adds vectors tip-to-tail?
  \item Resultant magnitude of 6 east and 8 north?
  \item Describe $-3\vec v$.
  \item Resultant magnitude of 7 east and 24 north?
  \item How do you express $\vec a-\vec b$ as an addition?
  \item Direction (N of E) of the resultant of 5 east, 5 north?
  \item Resultant magnitude of 9 east and 40 north?
  \item Length of $5\vec v$ if $|\vec v|=3$?
  \item Two forces, 10 N east and 24 N north, act on an object. Find the magnitude and direction of the resultant.
\end{enumerate}
\end{multicols}
\end{recapbox}

\begin{keybox}
\begin{multicols}{2}
\small
\begin{enumerate}[leftmargin=*,itemsep=3pt]
  \item $13$
  \item 4× length, same direction
  \item $\tfrac13$ length, opposite
  \item $17$
  \item triangle law
  \item $10$
  \item 3× length, opposite direction
  \item $25$
  \item $\vec a+(-\vec b)$
  \item $45^\circ$
  \item $41$
  \item $15$
  \item $\sqrt{100+576}=26$ N at $\tan^{-1}\tfrac{24}{10}\approx67^\circ$ N of E
\end{enumerate}
\end{multicols}
\end{keybox}

\end{document}
