\documentclass[11pt]{article}
\input{_preamble}
\fancyhead[R]{\footnotesize MCV4U \textperiodcentered{} 6.5}
\begin{document}

\begin{center}
{\LARGE\bfseries\color{titleblue}Worksheet 6.5 --- The Cross Product}\\[2pt]
{\small Grade 12 Calculus and Vectors (MCV4U) \textperiodcentered{} Unit 6: Geometry and Algebra of Vectors}
\end{center}

\begin{ideabox}
$\vec a\times\vec b=(a_2b_3-a_3b_2,\,a_3b_1-a_1b_3,\,a_1b_2-a_2b_1)$ is perpendicular to both; $|\vec a\times\vec b|$ is the parallelogram area.
\begin{itemize}[leftmargin=*,itemsep=2pt]
  \item $\vec a\times\vec b=(a_2b_3-a_3b_2,\,a_3b_1-a_1b_3,\,a_1b_2-a_2b_1)$.
  \item $\vec a\times\vec b\perp\vec a$ and $\perp\vec b$.
  \item $|\vec a\times\vec b|=$ parallelogram area; triangle $=\tfrac12|\vec a\times\vec b|$.
\end{itemize}
\end{ideabox}

\begin{learnbox}
\lh{Multiplying to a vector}The \textbf{cross product} $\vec u\times\vec v$ produces a \emph{vector} perpendicular to both $\vec u$ and $\vec v$ in 3-D.
\lh{Magnitude and area}Its magnitude $|\vec u|\,|\vec v|\sin\theta$ equals the area of the parallelogram the two vectors span.
\lh{Direction}The right-hand rule fixes the direction, and $\vec u\times\vec v=-\,\vec v\times\vec u$ — order matters.
\end{learnbox}

\sech{Worked Examples}

\begin{example}{ex0}{Example 1 --- Standard basis}
Compute $(1,0,0)\times(0,1,0)$.\soln $(0,0,1)$.
\end{example}

\begin{example}{ex1}{Example 2 --- General}
Compute $(1,2,3)\times(4,5,6)$.\soln $(2\cdot6-3\cdot5,\,3\cdot4-1\cdot6,\,1\cdot5-2\cdot4)=(-3,6,-3)$.
\end{example}

\begin{example}{ex2}{Example 3 --- Perpendicular check}
Verify $(-3,6,-3)\perp(1,2,3)$.\soln $-3+12-9=0$: perpendicular.
\end{example}

\begin{example}{ex3}{Example 4 --- Parallelogram area}
Area from $(1,0,0)$ and $(0,2,0)$.\soln Cross $=(0,0,2)$, magnitude $2$: area $2$.
\end{example}

\begin{example}{ex4}{Example 5 --- Triangle area}
Triangle area from $(1,0,0)$ and $(0,2,0)$.\soln $\tfrac12(2)=1$.
\end{example}

\begin{example}{ex5}{Example 6 --- Standard basis}
Compute $(0,1,0)\times(0,0,1)$.\soln $(1,0,0)$.
\end{example}

\begin{example}{ex6}{Example 7 --- General}
Compute $(2,0,0)\times(0,3,0)$.\soln $(0,0,6)$.
\end{example}

\begin{example}{ex7}{Example 8 --- General}
Compute $(1,1,0)\times(0,1,1)$.\soln $(1,-1,1)$.
\end{example}

\begin{example}{ex8}{Example 9 --- Parallelogram area}
Area from $(3,0,0)$ and $(0,4,0)$.\soln $(0,0,12)$, area $12$.
\end{example}

\clearpage
\sech{Your Turn --- Show Your Work}
For each question, show all steps clearly in the space provided.

\begin{qbox}{Question 1}
Compute $(0,1,0)\times(0,0,1)$.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 2}
Compute $(2,0,0)\times(0,3,0)$.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 3}
What direction is $\vec a\times\vec b$ relative to $\vec a,\vec b$?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 4}
Area of the parallelogram from $(3,0,0)$ and $(0,4,0)$?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 5}
Compute $(1,1,0)\times(0,1,1)$.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 6}
Compute $(1,0,0)\times(0,0,1)$.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 7}
Compute $(1,2,3)\times(4,5,6)$.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 8}
Triangle area from $(2,0,0)$ and $(0,2,0)$?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 9}
Is $(1,2,3)\times(2,4,6)$ the zero vector? Why?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 10}
Compute $(0,0,1)\times(1,0,0)$.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 11}
Magnitude of $(1,0,0)\times(0,5,0)$?\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 12}
Verify $(1,2,3)\times(4,5,6)=(-3,6,-3)$ is perpendicular to $(4,5,6)$.\workspace{2.4cm}
\end{qbox}

\begin{qbox}{Question 13 --- Challenge}
Find the area of the triangle with vertices $A(0,0,0)$, $B(1,0,0)$, $C(0,2,3)$.\workspace{3.5cm}
\end{qbox}

\clearpage
\sech{Question Recap \& Answer Key}

\begin{recapbox}
\begin{multicols}{2}
\small
\begin{enumerate}[leftmargin=*,itemsep=3pt]
  \item Compute $(0,1,0)\times(0,0,1)$.
  \item Compute $(2,0,0)\times(0,3,0)$.
  \item What direction is $\vec a\times\vec b$ relative to $\vec a,\vec b$?
  \item Area of the parallelogram from $(3,0,0)$ and $(0,4,0)$?
  \item Compute $(1,1,0)\times(0,1,1)$.
  \item Compute $(1,0,0)\times(0,0,1)$.
  \item Compute $(1,2,3)\times(4,5,6)$.
  \item Triangle area from $(2,0,0)$ and $(0,2,0)$?
  \item Is $(1,2,3)\times(2,4,6)$ the zero vector? Why?
  \item Compute $(0,0,1)\times(1,0,0)$.
  \item Magnitude of $(1,0,0)\times(0,5,0)$?
  \item Verify $(1,2,3)\times(4,5,6)=(-3,6,-3)$ is perpendicular to $(4,5,6)$.
  \item Find the area of the triangle with vertices $A(0,0,0)$, $B(1,0,0)$, $C(0,2,3)$.
\end{enumerate}
\end{multicols}
\end{recapbox}

\begin{keybox}
\begin{multicols}{2}
\small
\begin{enumerate}[leftmargin=*,itemsep=3pt]
  \item $(1,0,0)$
  \item $(0,0,6)$
  \item perpendicular to both
  \item $12$
  \item $(1,-1,1)$
  \item $(0,-1,0)$
  \item $(-3,6,-3)$
  \item $\tfrac12(4)=2$
  \item yes — they are parallel ($\sin\theta=0$)
  \item $(0,1,0)$
  \item $5$
  \item $-12+30-18=0$ ✓
  \item $\vec{AB}\times\vec{AC}=(1,0,0)\times(0,2,3)=(0,-3,2)$; area $=\tfrac12\sqrt{13}$
\end{enumerate}
\end{multicols}
\end{keybox}

\end{document}
